### 2021-12-10 smelt_2019_allseasons/f_clean_conditionfactor_survival.Rmd
Commit of clean data, code for cleaning data, code for plotting CF

---
title: "Clean_Condition_Survival"
author: "Catarina Pien"
date: "12/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(readr)
```

Tasks:
1. Read winter/spring, summer/fall data in
2. Clean names and bind together
3. Calculate Pre and Post condition factors
4. Make survival file

# Condition Factor, Length, Weight
1. Read data
```{r}
summerdata <- read_csv("smelt_2019_allseasons/Raw_Smelt-Length-Weight_Summer2019.csv", col_types = cols(Deployment = "f", Period = "f", Site = "f", ID = "c", FL_mm = "c", Weight_g= "d", Cage = "f", K =  "d", Comments =  "c")) %>%
  rename(K0 = K) 
falldata <- read_csv("smelt_2019_allseasons/Raw_Smelt-Length-Weight_Fall2019.csv", col_types = cols(Deployment = "f", Period = "f", Site = "f", ID = "c", FL_mm = "c", Weight_g= "d", Cage = "f", K =  "d", Comments =  "c")) %>%
  rename(K0 = K)

winterspringdata <- read_csv("smelt_2019_allseasons/Cagetype_data.csv") %>%
  rename(Pre_FL=Pre_FL_cm,
         Post_FL=Post_FL_cm,
         Pre_Weight=Pre_Weight_g,
         Post_Weight=Post_Weight_g)
```

2. Reformat
```{r}
winspring_long <- winterspringdata %>% 
  select(-Delta_CF, -X1) %>%
  pivot_longer(cols = c(Pre_FL:Post_CF),
               names_to = c("Period", ".value"),
               names_sep = "_")

# Filter to only large mesh and RV
winspring_clean <- winspring_long %>%
  filter(Site != "DWSC",
         Mesh == "large") %>%
  rename(FL_cm = FL,
         Weight_g = Weight,
         K0 = CF) %>%
  mutate(Deployment = ifelse(Site =="DWSC", "Spring",
                             ifelse(Site == "RV", "Winter", NA )),
         ID = NA,
         ID = as.character(ID),
         Deployment = as.factor(Deployment),
         Period = as.factor(Period),
         Site = as.factor(Site),
         Cage = as.factor(Cage))%>%
  select(Deployment, Period, Site, ID, FL_cm, Weight_g, Cage)

summerfall_clean <- rbind(summerdata, falldata) %>%
  mutate(FL_mm = replace(FL_mm, FL_mm == "-", NA),
         FL_mm = as.numeric(FL_mm),
         FL_mm = FL_mm/10, na.rm = TRUE) %>%
  rename(FL_cm = FL_mm) %>%
  select(Deployment, Period, Site, ID, FL_cm, Weight_g, Cage)

```

Bind 
```{r}
CF_2019 <- rbind(winspring_clean, summerfall_clean) %>%
  mutate(K= round(100 * Weight_g/(FL_cm^3),3)) 
```

Write file
```{r}
write.csv(CF_2019, "smelt_2019_allseasons/data_clean/clean_smelt_cf_2019.csv", row.names = FALSE)
```

# Survival
```{r}
Survival <- data.frame(Deployment = factor(), Site = factor(), Survived = numeric(), Total = numeric())

```

Winter
RV: 60, 64, 64, 64, 62, 56 / 64

Spring
DWSC: 59, 60, 58, 60, 60, 58 / 60

Summer
Yolo: 0 / 60
RV: / 60

Fall
Yolo: 
RV:
SM:
'
```{r}
(152-21)/180 #RV
(161-5)/180 #RV
(177-5-3)/180 #SM
(161-2)/180

Survivors_bycage <- CF_2019 %>%
  filter(Period == "Post",
         !Cage %in% c("FCCL1", "FCCL2")) %>%
  group_by(Deployment, Site, Cage) %>%
  summarize(n=n()) %>%
  mutate(Total = ifelse(Site == "RV", 64, 
                        ifelse(Site == "FCCL", 30, 60))) %>%
  filter(!is.na(Cage))

Survivors_bydeployment <-  Survivors_bycage %>%
  group_by(Deployment, Site) %>%
  summarize(Survived = sum(n),
            TotalFish = sum(Total)) %>%
  mutate(Prop = Survived/TotalFish) %>%
  ungroup()

Survivors_final <- Survivors_bydeployment %>% add_row(Deployment = "Summer", Site ="Yolo", Survived = 0, TotalFish = 30, Prop = 0)

```

```{r}
Cage_survival_RV <- as.data.frame(matrix(0, ncol = 2, nrow = 384))
colnames(Cage_survival_RV) <- c("Cage", "Survived")

Cage_survival_RV$Cage <- c(rep("A", 64), rep("B", 64), rep("C", 64), rep("D", 64), rep("E", 64), rep("F", 64))
Cage_survival_RV$Survived <- c(rep("1", 60), rep("0", 4), 
                               rep("1", 64), rep("0", 0),
                               rep("1", 64), rep("0", 0),
                               rep("1", 64), rep("0", 0),
                               rep("1", 62), rep("0", 2),
                               rep("1", 56), rep("0", 8))
```