### 2021-12-10 smelt_2019_allseasons/plot_weight_length_cf.Rmd

---
title: "plot_weight_length_cf"
author: "Catarina Pien"
date: "12/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
smeltdata <- read_csv("data_clean/clean_smelt_cf_2019.csv", col_types = cols(.default="f", FL_cm = "d", Weight_g = "d", K = "d", ID = "c"))
glimpse(smeltdata)
```

```{r}
smeltdata2 <- filter(smeltdata, Site !="DWSC")%>%
  filter(!is.na(K)) 
```

Summarize data
```{r}
summarydata <- smeltdata2 %>%
  group_by(Deployment, Site, Period) %>%
  summarize(meanCF = mean(K, na.rm = TRUE),
            sdCF = sd(K, na.rm = TRUE),
            meanFL = mean(FL_cm, na.rm = TRUE),
            sdFL = sd(FL_cm, na.rm = TRUE),
            meanWeight = mean(Weight_g, na.rm = TRUE),
            sdWeight = sd(Weight_g, na.rm = TRUE),
            n = n())
```

Plot data
```{r}
ggplot(smeltdata2, aes(x = Period, y = K, fill = Site)) + 
  geom_boxplot() + 
  facet_wrap(Deployment~.) + 
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw()
```