### 2022-02-28 smelt_2019_winterspring/check_length_distributions.Rmd
Commit of data and code previously written

---
title: "check_length-distribution"
author: "Catarina Pien"
date: "12/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(car)
library(FSA)
library(tidyverse)
library(ggthemes)
```

```{r}
Cagetype_data <- read.csv("smelt_2019_winterspring/Cagetype_data.csv")
Cagetype_data$Site <- ordered(Cagetype_data$Site, levels = c("RV", "DWSC"))
```

```{r}
NAs <- Cagetype_data %>%
  filter(is.na(Delta_CF),
         Mesh!= "Control")

```