### 2022-09-21 smelt_2019_winterspring/tag_loss.Rmd

---
title: "tag_loss"
author: "Catarina Pien"
date: '2022-09-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
DWSC_raw <- read.csv(here("smelt_2019_winterspring", "data_raw", "DWSC_Edited_Pre-Post_Data.csv"))%>%
  filter(!(grepl("FC", Cage))) 
RV_raw <- read.csv(here("smelt_2019_winterspring", "data_raw", "RVERS_Edited_Pre-Post_Data.csv")) %>%
  filter(!(grepl("FC", Cage)))

```

Look at no tags
```{r}
NT_DWSC <- DWSC_raw %>%
  mutate(NT = ifelse(grepl("NT", Tag), "NT", "Tag")) %>%
  mutate(Site = "DWSC")
NT_RV <- RV_raw %>%
  mutate(NT = ifelse(grepl("NT", Tag), "NT", "Tag")) %>%
  mutate(Site = "RV")
NT <- rbind(NT_DWSC, NT_RV) %>%
  mutate(Site = factor(Site, levels = c("RV", "DWSC")),
         Enclosure = ifelse(Cage %in% c("B","F"), "large", ifelse(Cage %in% c("A", "E" ), "wrap", ifelse(Cage %in% c("C", "D"), "small", "FC"))))

```

Tag loss rates - used these numbers in manuscript
```{r}
NT_prop <- NT %>%
  group_by(Site, Enclosure, NT) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = "NT", values_from = "n") %>%
  mutate(prop = NT/(NT+Tag)*100)
```

Plot
```{r}
NT_summary <- NT %>%
  filter(grepl("NT", Tag)) %>%
  group_by(Site, Enclosure) %>%
  summarize(n = n())

ggplot(NT_summary) + geom_col(aes(x = Enclosure,  y = n, fill = Site), position = "dodge2") + theme_bw()
```