### 2022-09-21 smelt_2019_winterspring/ZoopPlotting.Rmd
adding zoop data and analysis to project

---
librarytitle: "Zoop Plotting"
output: html_document
editor_options: 
  chunk_output_type: console
---

# load packages
```{r}
library(tidyverse)
library(lubridate)
library(emmeans)
library(multcomp)
setwd("C:/Users/nkwan/Desktop/Zoop'19")
```

# load data
```{r}
zoop <- read_csv("ZoopDataCombined_biomass_qc'ed_12.19.19.csv")

zoop$Date<-mdy(zoop$Date)

zoop$RingSize_m <- zoop$RingSize_cm/100

split<- split(zoop, zoop$Method)

write_excel_csv(split[["Throw"]], "C:/Users/nkwan/Desktop/Zoop'19/throw_method_zoop.csv")

write_excel_csv(split[["Tow"]], "C:/Users/nkwan/Desktop/Zoop'19/tow_method_zoop.csv")

throw<- read_csv("throw_method_zoop.csv")
tow<- read_csv("tow_method_zoop.csv")
```

# calculate abundance
```{r}
#calculate abundance for tow samples

tow$distance<-(tow$Rotations)*26873/999999 #R: Rotor constant (Regular flow = 26873, Low flow = 57560) - used Reg flow meter for this study
tow$volumesampled<-((pi*(tow$RingSize_m^2))/4)*tow$distance
tow$samplecount<-(tow$TotalVolume_ml/tow$Subsampled_ml)*tow$Count
tow$ind_m3<-tow$samplecount/tow$volumesampled

#to incorporate biomass
tow$samplebiomass<-(tow$TotalVolume_ml/tow$Subsampled_ml)*tow$Total.Biomass
tow$bio_m3<-tow$samplebiomass/tow$volumesampled

#ind_m3 = individuals/m3

#calculate abundance for throw samples

throw$distance<- "NA"
throw$radius<-throw$RingSize_m/2
throw$volumesampled<- throw$radius*2*pi*throw$TowDistance_m+(2*pi*(throw$radius^2))
throw$samplecount<-(throw$TotalVolume_ml/throw$Subsampled_ml)*throw$Count
throw$ind_m3<-throw$samplecount/throw$volumesampled

#to incorporate biomass
throw$samplebiomass<-(throw$TotalVolume_ml/throw$Subsampled_ml)*throw$Total.Biomass
throw$bio_m3<-throw$samplebiomass/throw$volumesampled
```

# bind throw and tow methods & calcualte species abundance for Table 6
```{r}
#combine
tow$radius<-tow$RingSize_m*2
abundance <- rbind(tow, throw)
abundance$Date<-ymd(abundance$Date)
abundance$Taxa<-as.factor(abundance$Taxa)
abundance$ind_m3<-as.numeric(abundance$ind_m3)

#put together results for Table 6
top.sp<-abundance %>% 
  group_by(SiteName, Taxa, ID) %>% 
  summarize("ind"= sum(ind_m3), "bio"=sum(bio_m3))

#remove microzooplankton
top.sp <- top.sp[top.sp$Taxa != "Microzoo",]

#sum by site (to get denominator)
sum<-top.sp %>% 
  group_by(SiteName) %>% 
  summarize("sum"= sum(bio))

#break into DWSC and RV
dw <- top.sp %>% filter(SiteName == "DWSC")
rv <- top.sp %>% filter(SiteName == "RV")

dw$DWSC.Bio <- dw$bio
dw$DWSC.Perc <- (dw$DWSC.Bio/12832.26)*100
dw <- dw[c(2,3,7,8)]
dw2 <- subset(dw, DWSC.Bio!=0)

rv$RV.Bio <- rv$bio
rv$RV.Perc <- (rv$RV.Bio/32782.81)*100
rv <- rv[c(2,3,7,8)]
rv2 <- subset(rv, RV.Bio!=0)

#merge
all <- merge(rv2,dw2,by=c("ID","Taxa"),all=TRUE)

#change NA to 0  
all[is.na(all)] <- 0

write_excel_csv(all, "species.table.csv")
```

# stacked bar across week plot
```{r}
taxa<-abundance %>% 
  group_by(SiteName, Date, Taxa) %>% 
  summarize("ind"= sum(ind_m3), "bio"=sum(bio_m3))

#create week column
taxa$Week <- epiweek(taxa$Date)

#zeros already exist, go ahead and average 
taxa.week<-taxa %>% 
  group_by(SiteName, Week, Taxa) %>% 
  summarize("bio.mean"=mean(bio))

taxa.week$Week2 <- NA

taxa.week$Week2[taxa.week$Week == 9] <- 1
taxa.week$Week2[taxa.week$Week == 10] <- 2
taxa.week$Week2[taxa.week$Week == 11] <- 3
taxa.week$Week2[taxa.week$Week == 12] <- 4
taxa.week$Week2[taxa.week$Week == 13] <- 5
taxa.week$Week2[taxa.week$Week == 4] <- 1
taxa.week$Week2[taxa.week$Week == 5] <- 2
taxa.week$Week2[taxa.week$Week == 6] <- 3
taxa.week$Week2[taxa.week$Week == 7] <- 4
taxa.week$Week2[taxa.week$Week == 8] <- 5

#plot all zoop

cbPalette <- c("#009E73", "#E69F00", "#56B4E9", "#D55E00", "#F0E442", "#999999", "#CC79A7")
legend_title <- "Taxa"

ggplot(data = taxa.week, aes(x=Week2, y=bio.mean)) +
  geom_bar(stat = "identity", data=taxa.week, aes (x=Week2, y=bio.mean, fill=factor(Taxa, levels=c("Cladocera", "Calanoid", "Cyclopoid", "Harpacticoid", "Macrozoo", "Microzoo"))), colour="black", width = .5) +
  facet_wrap(~SiteName)+
  theme_bw() + 
    scale_fill_manual(legend_title, values=cbPalette)+theme(axis.title.x = element_blank()) +
  labs(x = "Week") +
  labs(y = "Biomass (ugC/m3)")+
 theme(legend.position="bottom")+theme(axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25))

#remove microzooplankton
taxa.week.2 <- taxa.week[taxa.week$Taxa != "Microzoo",]

taxa.week.2$SiteName = factor(taxa.week.2$SiteName, levels=c('RV','DWSC'))

taxa.week.2$Taxa <- as.character(taxa.week.2$Taxa)
taxa.week.2$Taxa[taxa.week.2$Taxa=='Macrozoo'] <- 'Other'

windowsFonts(Times = windowsFont("Times New Roman"))

#plot just macrozoop
z<-ggplot(data = taxa.week.2, aes(x=Week2, y=bio.mean)) +
  geom_bar(stat = "identity", data=taxa.week.2, aes (x=Week2, y=bio.mean, fill=factor(Taxa, levels=c("Cladocera", "Calanoid", "Cyclopoid", "Harpacticoid", "Other"))), colour="black", width = .5) +
    facet_wrap(~SiteName)+
  theme_bw() + 
    scale_fill_manual(legend_title, values=cbPalette)+
  labs(x = "Week") +
  labs(y = "Biomass (ugC/m3)")+
 theme(legend.position="right")+
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size =17),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 17),
        strip.text.x = element_text(size = 17),
        text=element_text(family = "Times"))
z
print(z)

tiff(filename="Fig5.tiff", 
    units="in", 
    width=7.25, #or 3.5 for single column
    height=3.75, #height you can adjust according to your plot
    res=600, compression="lzw")
print(z)
dev.off()
```

# repeated measures ANOVA
```{r}
#combine taxa 
for.mod<-taxa.week.2 %>% 
  group_by(SiteName, Week, Week2) %>% 
  summarize("bio.all"=sum(bio.mean))

shapiro.test(for.mod$bio.all) #not significant

model<-lm(bio.all~SiteName*Week2, data=for.mod)
anova(model)

shapiro.test(resid(model)) #insigificant

leveneobject<- lm(abs(resid(model))~SiteName*Week2, data=for.mod)
anova(leveneobject) #insignificant

marginal = emmeans(model, 
                   ~ SiteName*Week2)

cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     
    adjust  = "tukey") #DWSC and RV different
```