### 2022-09-22 smelt_2019_winterspring/Diet.Rmd
updates to diet analysis to explore random effect of cage

---
editor_options: 
  chunk_output_type: console
---

# load packages
```{r}
library(tidyverse)
library(lubridate)
library(emmeans)
library(multcomp)
library(rcompanion)
library(lme4)
library(Rcpp)
```

# bring in data
```{r}
diet <- read_csv("smelt_2019_winterspring/data_raw/DSM Cage Diets 2019 update_taxa.csv")
diet$Location <- ordered(diet$Location, levels = c("RVR", "DWSC"))
diet<-diet[diet$Cage.ID != "FCCL Control Tank 1",]
diet<-diet[diet$Cage.ID != "FCCL Control Tank 2",]
diet<-diet[diet$Cage.ID != "FCCL Control",]

#fix fish that were assigned the wrong cage (came up during summary stats, check against dissected fish data, likely E and F were misread)
diet[diet$Tag=='RK10', 'Cage.ID'] <- "E"
diet[diet$Tag=='RN93', 'Cage.ID'] <- "E"

#create new column for mesh types
Cage_type <- as.data.frame(matrix(0, ncol = 2, nrow = 6))
colnames(Cage_type) <- c("Cage.ID", "Mesh")

Cage_type$Cage.ID <- c("A", "B", "C", "D", "E", "F")
Cage_type$Mesh <- c("wrap", "large", "small", "small", "wrap", "large")

diet_type <- merge(diet, Cage_type, by = "Cage.ID", all.x = TRUE)

#pull in weight data
DWSC <- read.csv("smelt_2019_winterspring/data_raw/DWSC_Edited_Pre-Post_Data.csv")
DWSC$Site <- "DWSC"
RVR <- read.csv("smelt_2019_winterspring/data_raw/RVERS_Edited_Pre-Post_Data.csv")
RVR$Site <- "RVR"
Cage <- rbind(DWSC, RVR)

#bind fish measurements with stomach contents (note- some lost due to missing letters in ID's-update for final)
total <- merge(diet_type, Cage, by=c("Tag"))

#create instantaneous ration column
total$IR <- (total$Contents.Weight/total$Post_Weight_g)*100

#average for IR & fullness
mean.IR<- total %>% 
  group_by(Location, Mesh, Tag, Cage.ID) %>% 
  summarize("mean.IR"= mean(IR), "mean.full"=mean(Fullness))

#split into RV and DWSC
split<- split(mean.IR, mean.IR$Location)
#write_excel_csv(split[["RVR"]], "smelt_2019_winterspring/data_raw/diet.RVR.csv")
#write_excel_csv(split[["DWSC"]], "smelt_2019_winterspring/data_raw/diet.DWSC.csv")
RVR<- read_csv("smelt_2019_winterspring/data_raw/diet.RVR.csv")
DWSC<- read_csv("smelt_2019_winterspring/data_raw/diet.DWSC.csv")

RVR$Mesh<-as.factor(RVR$Mesh)
DWSC$Mesh<-as.factor(DWSC$Mesh)
```

# stats on IR
```{r}
##Shapiro-Wilk's method
shapiro.test(RVR$mean.IR) #significant
shapiro.test(DWSC$mean.IR) #significant

#log transform
RVR$mean.IR.log <- log(RVR$mean.IR +1)
shapiro.test(RVR$mean.IR.log) #insignificant
plotNormalHistogram(RVR$mean.IR.log)

DWSC$mean.IR.log <- log(DWSC$mean.IR +1)
shapiro.test(DWSC$mean.IR.log) #insignificant
plotNormalHistogram(DWSC$mean.IR.log)

#RVR anova 
RVR.IR.mod <- lm(mean.IR.log ~ Mesh, data=RVR)
anova(RVR.IR.mod) #significant

shapiro.test(resid(RVR.IR.mod)) #insignificant

leveneobject<- lm(abs(resid(RVR.IR.mod))~Mesh, data=RVR)
anova(leveneobject) #good

#post hoc test
marginal = emmeans(RVR.IR.mod, 
                   ~ Mesh)

cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     
    adjust  = "tukey")

#Should Cage.ID be included as a random effect? Looked over with Rosie and lmer model gives singularity issue, seperate models with Cage.ID don't show significance - okay to leave out
  #ggplot(RVR,aes(x=Mesh,y=mean.IR.log,col=Cage.ID)) + geom_jitter() + geom_boxplot(alpha=0.2)
  #RVR.IR.mod2 <- lmer(mean.IR.log ~ Mesh + (1|Cage.ID), data=RVR)
  #summary(mod)
  #ranef(RVR.IR.mod2) #saying no variance by cage
  #RVR.IR.mod3 <- lm(mean.IR.log ~ Cage.ID, data=RVR)
  #RVR.IR.mod4 <- lm(mean.IR.log ~ Mesh + Cage.ID, data=RVR)

#DWSC anova 
DW.IR.mod <- lm(mean.IR.log ~ Mesh, data=DWSC)
anova(DW.IR.mod) #insignificant

DW.IR.mod2 <- lmer(mean.IR.log ~ Mesh + (1|Cage.ID), data=DWSC)

ranef(DW.IR.mod2) #close to 0 but some variance

ggplot(DWSC,aes(x=Mesh,y=mean.IR.log,col=Cage.ID)) + geom_jitter() + geom_boxplot(alpha=0.2)

anova(DW.IR.mod2)

shapiro.test(resid(DW.IR.mod)) #insignificant

leveneobject<- lm(abs(resid(DW.IR.mod))~Mesh, data=DWSC)
anova(leveneobject) #good

#post hoc test
marginal = emmeans(DW.IR.mod, 
                   ~ Mesh)

cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     
    adjust  = "tukey")
```

# proportion taxa
```{r}
diet_type$Count<-as.numeric(diet_type$Count)

#sum for taxa abundance among individual fish
taxa_tag<- diet_type %>% 
  group_by(Location, Mesh, Cage.ID, Tag, Taxa, LH.Stage) %>% 
  summarize("count.sum"= sum(Count))

#mean for stomach mass
taxa<- taxa_tag %>% 
  group_by(Location, Mesh, Cage.ID, Tag) %>% 
  summarize("stomach.sum"= sum(count.sum))

stomach_abundance<-merge(taxa_tag, taxa, by = "Tag")

stomach_abundance$proportion<-(stomach_abundance$count.sum/stomach_abundance$stomach.sum)*100

#add rows for all taxa for all fish
full.prop<-stomach_abundance %>% complete(Tag, Taxa) %>%
  as.data.frame()

full.prop.2<-merge(full.prop, taxa, by = "Tag")

#replace count.sum Nas with zeros
full.count<-full.prop.2 %>% replace_na(list(count.sum=0))

#select columns I want for count
full.count<-subset(full.count, select=c("Location", "Mesh", "Tag", "Taxa", "LH.Stage", "count.sum"))

#select columns I want for proportion
full.prop.3<-subset(full.prop.2, select=c("Tag", "Taxa", "proportion", "Location", "Mesh"))
#replace proportion Nas with zeros
diet.prop<-full.prop.3 %>% replace_na(list(proportion=0))

#average across fish
diet.prop.mean<- diet.prop %>% 
  group_by(Taxa, Location, Mesh) %>% 
  summarize("prop.mean"= mean(proportion))
```

# plot mean proportion in bar plot and 100% stacked bar
```{r}
#remove empty
diet.prop.mean.e<-diet.prop.mean[diet.prop.mean$Taxa != "empty",]

#change labels
diet.prop.mean.e$Mesh[diet.prop.mean.e$Mesh == "large"] <- "B"
diet.prop.mean.e$Mesh[diet.prop.mean.e$Mesh == "small"] <- "A"
diet.prop.mean.e$Mesh[diet.prop.mean.e$Mesh == "wrap"] <- "C"

#plot
windowsFonts(Times = windowsFont("Times New Roman"))

d<-ggplot(data=diet.prop.mean.e, aes(x=Mesh, y =prop.mean)) +
    geom_bar(aes(fill=factor(Taxa, levels=c("Amphipoda", "Calanoida", "Cladocera", "Cyclopoida", "Diptera", "Harpacticoda", "Ostracoda", "Other"))), position="fill", stat="identity") +
  facet_grid(~Location)+
  theme_bw() + 
  scale_fill_manual(legend_title, values=cbPalette)+
  scale_y_continuous(labels = scales::percent)+
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Mean Contribution to Diet") +  
  theme(panel.grid.major = element_line(colour = "gray93",size=0.01), panel.grid.minor = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.title = element_text(size = 17), 
    axis.text = element_text(size = 16), 
    legend.text = element_text(size = 16), 
    legend.title = element_text(size = 17),
    strip.text.x = element_text(size = 17),
    text=element_text(family = "Times"))
d

tiff(filename="smelt_2019_winterspring/figures/dietfig.tiff", 
    units="in", 
    width=7.25, #or 3.5 for single column
    height=4.00, #height you can adjust according to your plot
    res=600, compression="lzw")
print(d)
dev.off()
```

# percent of each taxa for potential supplemental table
```{r}
diet_type$Count[which(diet_type$Count == "1 maybe 2")] <- 2

diet_type$Count <- as.numeric(diet_type$Count)

taxa.perct<- diet_type %>% 
  group_by(Location, Taxa, Prey.Taxa) %>% 
  summarize("count.sum"= sum(Count, na.rm=TRUE))

taxa.tot<- diet_type %>% 
  group_by(Location) %>% 
  summarize("count.sum"= sum(Count, na.rm=TRUE))

taxa.perct$Total <- ifelse(taxa.perct$Location=="RVR", 1548,
        ifelse(taxa.perct$Location=="DWSC", 5831, NA))

taxa.perct$percent <- (taxa.perct$count.sum/taxa.perct$Total)*100
```