### 2022-10-18 smelt_2019_allseasons/Zoop 2019.Rmd

---
librarytitle: "Zoop Plotting"
output: html_document
editor_options: 
  chunk_output_type: console
---
# load packages
```{r}
library(tidyverse)
library(lubridate)
library(Polychrome)
```

# load & prep summer/fall data
```{r}
zoop <- read_csv("hybrid study/ZoopData_SummerFall_2019.csv")

zoop$Date <- mdy(zoop$Date)
zoop$Day <- day(zoop$Date)
zoop$Month <- month(zoop$Date)
zoop$Month <- month.abb[zoop$Month]
zoop$MD <- paste(zoop$Month, zoop$Day, sep="-")

#add ring size (0.5m for 50um net)
zoop$RingSize_m <- 0.5

#add 0s for count with NA
zoop$Count[is.na(zoop$Count)] <- 0

#bring in tow data
aug <- read_csv("hybrid study/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
oct <- read_csv("hybrid study/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")
aug$Date <- mdy(aug$Date)
oct$Date <- mdy(oct$Date)

#condense to columns of interest
aug2 <- aug[c(1,4,22,24:27)]
oct2 <- oct[c(1,4,26,28:31)]

tow <- rbind(aug2,oct2)
tow2 <- tow[!is.na(tow$Location),]

#merge
#rename column
zoop <- zoop %>% rename(Location = "Station")

#make locations match
unique(zoop$Location)
unique(tow2$Location)

zoop[zoop$Location %in% c("STTD", "STTD-YOLO", "YOLO"),]$Location <- "Yolo"

all <- merge(zoop, tow2, by = c('Location', 'Date'))

#calculate abundance -----------------------
all$distance <- (all$Revs)*57560/999999
all$volumesampled <- ((pi*(all$RingSize_m^2))/4)*all$distance
all$samplecount <- (all$TotalVolume_ml/all$Subsampled_ml)*all$Count

#calculate ind/m3
all$ind_m3 <- all$samplecount/all$volumesampled

#remove 0's
caught <- all[!(all$ind_m3==0),]
```
# plot summer/fall data
```{r}
#order factors
unique(caught$MD)
caught$MD <- factor(caught$MD, levels=c("Jul-31", "Aug-1", "Aug-2",  "Aug-5", "Aug-6", "Aug-8",  "Aug-12", "Aug-15", "Aug-19", "Aug-22", "Aug-26", "Oct-10", "Oct-11", "Oct-15", "Oct-16", "Oct-18", "Oct-21", "Oct-25", "Oct-28", "Nov-4"))

legend_title <- "Species"

pal36 <- palette36.colors(36)
pal36 <- as.vector(t(pal36))

ggplot(data = caught, aes(x=MD, y=samplecount)) +
  geom_bar(position = "fill", stat = "identity", data=caught, aes (x=MD, y=samplecount, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of all Zoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

ggsave("hybrid study/plots/allzoop.png", dpi=250, height=6, width=10, units="in")

#plot with microzoop removed
macro <- caught[!(caught$ID=="Copepod nauplii"),]
macro <- macro[!(macro$ID=="Rotifers"),]
macro <- macro[!(macro$ID=="Ostracods"),]

ggplot(data = macro, aes(x=MD, y=samplecount)) +
  geom_bar(position = "fill", stat = "identity", data=macro, aes (x=MD, y=samplecount, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of Macrozoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

ggsave("hybrid study/plots/macrozoop.png", dpi=250, height=6, width=10, units="in")
```
#bring in winter data
```{r}
#prep of this dataset (code and raw data) can be found here: https://cawater.sharepoint.com/:f:/r/sites/dwrdoc-aes/Shared%20Documents/Delta%20Smelt%20Enclosure%20Study/Analysis/Zooplankton/Zoop%2719?csf=1&web=1&e=YhKzMy

winter <- read_csv("hybrid study/zoop_abundance.csv")

winter$Day <- day(winter$Date)
winter$Month <- month(winter$Date)
winter$Month <- month.abb[winter$Month]
winter$MD <- paste(winter$Month, winter$Day, sep="-")

#remove throw samples
winter <- winter[!(winter$Method=="Throw"),]

#remove DWSC samples
winter <- winter[!(winter$SiteName=="DWSC"),]

#merge with catch df

#condense to columns of interest
winter2 <- winter[c(1,2,13,16,32,25,26)]
caught2 <- caught[c(1,2,7,10,20,29,30)]

#rename column
winter2 <- winter2 %>% rename(Location = "SiteName")

#make locations match
unique(winter2$Location)
unique(caught2$Location)

all2019 <- rbind(winter2, caught2)

#remove 0's
caught.19 <- all2019[!(all2019$ind_m3==0),]
```
# plot all 2019
```{r}
#order factors
unique(caught.19$MD)
caught.19$MD <- factor(caught.19$MD, levels=c("Jan-24", "Jan-29", "Jan-31", "Feb-6", "Feb-8", "Feb-12", "Feb-15", "Feb-22", "Jul-31", "Aug-1", "Aug-2",  "Aug-5", "Aug-6", "Aug-8",  "Aug-12", "Aug-15", "Aug-19", "Aug-22", "Aug-26", "Oct-10", "Oct-11", "Oct-15", "Oct-16", "Oct-18", "Oct-21", "Oct-25", "Oct-28", "Nov-4"))

ggplot(data = caught.19, aes(x=MD, y=samplecount)) +
  geom_bar(position = "fill", stat = "identity", data=caught.19, aes (x=MD, y=samplecount, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of all Zoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

ggsave("hybrid study/plots/allzoop19.png", dpi=250, height=6, width=10, units="in")

#plot with microzoop removed
macro2 <- caught.19[!(caught.19$ID=="Copepod nauplii"),]
macro2 <- macro2[!(macro2$ID=="Rotifers"),]
macro2 <- macro2[!(macro2$ID=="Ostracods"),]

ggplot(data = macro2, aes(x=MD, y=samplecount)) +
  geom_bar(position = "fill", stat = "identity", data=macro2, aes (x=MD, y=samplecount, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of Macrozoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

ggsave("hybrid study/plots/macrozoop19.png", dpi=250, height=6, width=10, units="in")
```