### 2022-11-19 smelt_2019_allseasons/Diet 2019.Rmd
updating diet plots

---
title: "Delta Smelt Enclosure Study 2019 Hybrid Paper Diet Analysis"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

# load packages
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(FSA)
library(lubridate)
```

# bring in contractor (UW) data
```{r}
diet2 <- read_csv("smelt_2019_allseasons/data_raw/DSM Cage Diets 2019-2020.csv") #new data, not edited at all

diet2$Date <- mdy(diet2$Date)

#format 
#rename columns
diet2 <- diet2 %>% rename(Cage.ID = "Cage ID", Contents.Weight = "Total Contents Weight", Prey.Taxa = "Prey Taxa", LH.Stage = "LH Stage")

#add taxa column (choose taxa based on what is present in the dataframe)
diet2$Taxa <- as.character(NA)

str(diet2)

diet2[diet2$Prey.Taxa %in% c("Bosmina sp.", "Chydorus sp.", "Daphnia sp.", "Ceriodaphnia sp.", "Chydoridae", "Crangonyx sp.", "Scapholeberis sp.", "Simocephalus sp.", "Eurycercus sp.", "Leydigia sp.", "Cladocera", "Sida sp.", "Alona sp.", "Diaphanosoma sp."),]$Taxa <- "Cladocera"

diet2[diet2$Prey.Taxa %in% c("Americorophium sp.", "Americorophium spinicorne", "Eogammarus confervicolus", "Gammaridea", "Hyallela azteca", "Crangonyx sp.", "Corophiidae", "Amphipoda"),]$Taxa <- "Amphipod"

diet2[diet2$Prey.Taxa %in% c("Pseudodiaptomus forbesi", "Calanoida", "Eurytemora affinis", "Diaptomidae", "Sinocalanus doerrii", "Osphranticum labronectum"),]$Taxa <- "Calanoid"

diet2[diet2$Prey.Taxa %in% c("Acanthocyclops sp.", "Macrocyclops sp.", "Diacyclops thomasi", "Eucyclops sp.", "Cyclopoida", "Halicyclops sp.", "Mesocyclops sp.", "Tropocyclops sp.", "Tropocyclops sp."),]$Taxa <- "Cyclopoid"

diet2[diet2$Prey.Taxa %in% c("Chironomidae", "Diptera"),]$Taxa <- "Diptera"

diet2[diet2$Prey.Taxa %in% c("empty", "EMPTY"),]$Taxa <- "empty"

diet2[diet2$Prey.Taxa %in% c("Harpacticoida"),]$Taxa <- "Harpacticoid"

diet2[diet2$Prey.Taxa %in% c("Ostracoda"),]$Taxa <- "Ostracod"

diet2[diet2$Prey.Taxa %in% c("Fish", "Acari", "Insecta", "Bivalvia", "Culicidae", "egg unidentified", "Corixidae", "Hemiptera", "plant seed", "plant matter", "Unidentified", "Mysidacea", "Cumacea"),]$Taxa <- "Other"

new_DF <- diet2[is.na(diet2$Taxa),] #update code above with new prey.taxa that weren't found in previous diets
#there are two "copepoda" entries with no count or location...don't seem that important so I am going to just go ahead and remove them
diet2[diet2$Taxa %in% c("empty"),]$Count <- "0"
diet2 <- diet2[!is.na(diet2$Count),]

#Tag column currently has rows with just the Tag and others with location, date, and tag
#first break those two kinds of columns apart
weird.tag <- diet2[is.na(diet2$Location),]
norm.tag <- diet2[!is.na(diet2$Location),]

#then for break out the Tag ID into Location and tag (remove date)
weird.tag <- weird.tag %>% separate(Tag, c("Location", "X.date", "Tag"))

#remove new date column
weird.tag <- subset(weird.tag, select = -6)

#add mesh types for prototype experiment
#create new column for mesh types
Cage_type <- as.data.frame(matrix(0, ncol = 2, nrow = 6))
colnames(Cage_type) <- c("Cage.ID", "Mesh")

Cage_type$Cage.ID <- c("A", "B", "C", "D", "E", "F")
Cage_type$Mesh <- c("small", "large", "wrap", "wrap", "small", "large")

norm.tag2 <- merge(norm.tag, Cage_type, by = "Cage.ID", all.x = TRUE)

#add mesh types (NA) for summer/fall experiements
weird.tag$Mesh <- NA

#merge the two
diet <- rbind(weird.tag, norm.tag2)

#change location titles 
diet$Location[diet$Location =="RIVERS"] <- "RV.Winter"
diet$Location[diet$Location =="Sac DWSC"] <- "DWSC.Winter"
diet$Location[diet$Location =="SM"] <- "SM.Fall"
diet$Location[diet$Location =="YB"] <- "YB.Fall"
#RV a little trickier with two dates
diet$ld = paste(diet$Location, diet$Date)
diet$Location[diet$ld =="RV 2019-08-28"] <- "RV.Summer"
diet$Location[diet$ld =="RV 2019-11-06"] <- "RV.Fall"
RV <- filter(diet, Location == "RV")
diet$Location[diet$ld =="RV 2019-11-07"] <- "RV.Fall"

diet <- subset(diet, select = -16)
```

# prep merged data for taxa proportion
```{r}
#remove FCCL controls
diet<-diet[diet$Cage.ID != "FCCL Control Tank 1",]
diet<-diet[diet$Cage.ID != "FCCL Control Tank 2",]
diet<-diet[diet$Cage.ID != "FCCL Control",]

#fix fish that were assigned the wrong cage (came up during summary stats, check against dissected fish data, likely E and F were misread)
diet[diet$Tag=='RK10', 'Cage.ID'] <- "E"
diet[diet$Tag=='RN93', 'Cage.ID'] <- "E"

#change count to numeric
diet$Count<-as.numeric(diet$Count)

#output says NAs introduced by coercion - check NA's
newDF <- diet[is.na(diet$Count),]

#one NA next to a gammaridae - comment says that it was too digested and could be more than one individual, to be safe just going to add a count of "1" for this observation; found it on row 413
diet<-diet %>% replace_na(list(Count=1))

#change count to numeric
diet$Count<-as.numeric(diet$Count)

diet$tlc = paste(diet$Tag, diet$Location, diet$Cage.ID)

#sum for taxa abundance among individual fish
taxa_tag<- diet %>% 
  group_by(Location, Mesh, tlc, Taxa) %>% 
  summarize("count.sum"= sum(Count))

#sum taxa across stomach
stomach<- diet %>% 
  group_by(Location, Mesh, tlc) %>% 
  summarize("stomach.sum"= sum(Count))

stomach_abundance<-merge(taxa_tag, stomach, by = "tlc")

stomach_abundance$proportion<-(stomach_abundance$count.sum/stomach_abundance$stomach.sum)*100

#add rows for all taxa for all fish so every taxa shows up for every stomach
full.prop<-stomach_abundance %>% complete(tlc, Taxa) %>%
  as.data.frame()

full.prop.2<-merge(full.prop, stomach, by = "tlc")

#replace count.sum Nas with zeros
full.count<-full.prop.2 %>% replace_na(list(count.sum=0))

#select columns I want for count
full.count<-subset(full.count, select=c("Location", "Mesh", "tlc", "Taxa", "count.sum"))

#select columns I want for proportion
full.prop.3<-subset(full.prop.2, select=c("Location", "Mesh", "tlc", "Taxa", "proportion"))

#replace proportion Nas with zeros
diet.prop<-full.prop.3 %>% replace_na(list(proportion=0))

#average across fish
diet.prop.mean<- diet.prop %>% 
  group_by(Taxa, Location) %>% 
  summarize("prop.mean"= mean(proportion))
```

# plot number of empty stomachs
## these will not be included in the 100% stacked bar later
```{r}
#order locations
stomach_abundance$Location <- stomach_abundance$Location.x
stomach_abundance$Location <- ordered(stomach_abundance$Location, levels = c("DWSC.Winter", "RV.Winter", "RV.Summer", "SM.Fall", "YB.Fall", "RV.Fall"))

#subset for only empty stomachs
empty <- filter(stomach_abundance, Taxa == "empty")

ggplot(empty, aes(Location)) + geom_bar() + scale_x_discrete(drop=FALSE)
```

# plot mean proportion in 100% stacked bar
```{r}
#remove empty
prop<-diet.prop.mean[diet.prop.mean$Taxa != "empty",]

#order locations
prop$Location <- ordered(prop$Location, levels = c("DWSC.Winter", "RV.Winter", "RV.Summer", "RV.Fall", "YB.Fall", "SM.Fall"))

#order taxa to match zoop plot
prop$Taxa <- factor(prop$Taxa, levels=c("Calanoid", "Cladocera", "Cyclopoid", "Harpacticoid", "Ostracod", "Other", "Amphipod", "Diptera"))

ggplot(data=prop, aes(x=Location, y =prop.mean, fill=Taxa)) +
    geom_bar(position="fill", stat="identity", colour="black") +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Mean Contribution to Diet")

ggsave("smelt_2019_allseasons/figures/diet19_onepanel.png", dpi=250, height=4.5, width=7, units="in")

#facet by season

#create season column
prop2 <- prop
prop2$Season <- "NA"
prop2[prop2$Location %in% c("DWSC.Winter", "RV.Winter"),]$Season <- "Winter"
prop2[prop2$Location %in% c("RV.Summer"),]$Season <- "Summer"
prop2[prop2$Location %in% c("RV.Fall", "YB.Fall", "SM.Fall"),]$Season <- "Fall"

#change Location to drop season
prop2$Location <- as.character(prop2$Location)
prop2$Location[prop$Location =="DWSC.Winter"] <- "DWSC"
prop2[prop2$Location %in% c("RV.Winter", "RV.Summer", "RV.Fall"),]$Location <- "RV"
prop2$Location[prop$Location =="YB.Fall"] <- "YB"
prop2$Location[prop$Location =="SM.Fall"] <- "SM"

#order seasons
prop2$Season <- factor(prop2$Season, levels=c("Winter", "Summer", "Fall"))

#order locations
prop2$Location <- factor(prop2$Location, levels=c("DWSC", "RV", "YB", "SM"))

#order taxa to match zoop plot
prop2$Taxa <- factor(prop2$Taxa, levels=c("Calanoid", "Cladocera", "Cyclopoid", "Harpacticoid", "Ostracod", "Other", "Amphipod", "Diptera"))

#plot
ggplot(data=prop2, aes(x=Location, y =prop.mean, fill=Taxa)) +
  geom_bar(position="fill", stat="identity", colour="black", width=0.75) +
  facet_wrap(~Season, ncol=1) +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Mean Contribution to Diet") +
   theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10))

library(scales)
show_col(viridis_pal(option = "plasma")(8))
viridis_pal(option = "plasma")(8)

ggsave("smelt_2019_allseasons/figures/diet19_multipanel.png", dpi=250, height=6, width=6, units="in")
```

### Plot methods text (from manuscript): To visualize the relative abundance of different taxa among mesh types, we calculated taxa proportion (taxa abundance/total abundance) for each individual fish’s diet and then averaged these taxa proportions across individuals from each mesh type. These means were then fit to a 100% scale to estimate the mean percent contribution of each taxa to overall mesh type diets. Averaging proportions instead of abundance across fish allowed for the diets of different sizes fish to contribute equally.