### 2023-03-06 recaps.Rmd
added smelt recapture stats

---
title: "Recpature distances"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(readxl)
library(sf)
library(lubridate)
library(spacetools)
library(deltamapr)
load("Recaps.RData")
```

## Spatial stats on Delta Smelt releases

We have four different release types and a few questions we can ask.

1. How far did smelt from each release group travel before being recaptured?
2. How variable was the distance between release and recapture?
3. How close together were fish from a particular release when they were recaptured?
4. How variable was the distance between fish on recapture?

Sam Bashevkin has written some nifty functions for calculating in=water distance between points (fish) and a reference point (release location) or between a group of points (fish from a particular release).

https://github.com/sbashevkin/spacetools

Fist let's calcualte distance from release location to recapture location.

```{r}
#Bring in the data and check it out

recaps = read_excel("Recaptured Experimental Release_2022-2023_working draft.xlsx", 
                    sheet = "Adult Recaptures") %>%
  rename(Site = `Release Site`, Type = `Release Type`, Latitude= `Start Latitude`, Longitude = `Start Longitude`)
recaps    =     mutate(recaps, fishID = paste("smelt", 1:nrow(recaps)),
                Type = case_when(Type == "header" ~ "Soft",
                                 TRUE ~ Type))
str(recaps)

```

This just needs a little data manipulation

```{r}

recapsf = st_as_sf(filter(recaps, !is.na(`Longitude`)), coords = c("Longitude", "Latitude"), crs = 4326)

#Now make a data frame of the release locations
release = data.frame(Site = c("SDWSC", "SDWSC", "Rio Vista", "Rio Vista"), Type = c("Soft", "Hard", "Hard", "Hard (Trailer)"),
                     Latitude = c(38.237842, 38.275563, 38.124736, 38.139752), 
                     Longitude = c(-121.673802, -121.661374, -121.699521, -121.694706)) 
releasesf = st_as_sf(release,  coords = c("Longitude", "Latitude"), crs = 4326)

#merge recap points with release points

recaps2 = left_join(recaps, release)

#make a quick plot to make sure I did it right
ggplot()+
  geom_sf(data = WW_Delta)+
  geom_sf(data = recapsf, aes(color = Type))+
  geom_sf(data = releasesf, aes(color = Type), size = 5)+

  coord_sf(xlim = c(-121.5, -122.2), ylim = c(37.8, 38.5))

```

Now we can calculate the distance between each point and the release location. I'm going to do it one release at a time. THere is probalby a way to vectorize this and do it all at once, but I'm too lazy to figure it out.

```{r, eval=FALSE}
#one release at a time
Dwschard = filter(recaps2, Type == "Hard", Site == "SDWSC")
dwsch = filter(release, Site == "SDWSC", Type == "Hard") %>%
  select(Latitude, Longitude)
tesst = GGdist(Water_map = spacetools::Delta, Points = Dwschard, Latitude_column = `Latitude`,
               Longitude_column = `Longitude`, PointID_column = fishID, EndPoint = dwsch)

#now let's do DWSC soft
Dwscsoft = filter(recaps2, Type == "Soft", Site == "SDWSC")
dwscs = filter(release, Site == "SDWSC", Type == "Soft") %>%
  select(Latitude, Longitude)
tesst2 = GGdist(Water_map = spacetools::Delta, Points = Dwscsoft, Latitude_column = `Latitude`,
               Longitude_column = `Longitude`, PointID_column = fishID, EndPoint = dwscs)

#now let's do Rio Vista hard
RVhard = filter(recaps2, Type == "Hard", Site == "Rio Vista", !is.na(Latitude))
rvh = filter(release, Site == "Rio Vista", Type == "Hard") %>%
  select(Latitude, Longitude)
tesst3 = GGdist(Water_map = spacetools::Delta, Points = RVhard, Latitude_column = `Latitude`,
                Longitude_column = `Longitude`, PointID_column = fishID, EndPoint = rvh)

#now the trailer
RVhardt = filter(recaps2, Type == "Hard (Trailer)", Site == "Rio Vista")
rvht = filter(release, Site == "Rio Vista", Type == "Hard (Trailer)") %>%
  select(Latitude, Longitude)
tesst4 = GGdist(Water_map = spacetools::Delta, Points = RVhardt, Latitude_column = `Latitude`,
                Longitude_column = `Longitude`, PointID_column = fishID, EndPoint = rvht)

distances = bind_rows(tesst, tesst2, tesst3, tesst4)
```

OK, now let's bind all the distances together and join them back to the origional dataset

```{r}

recaps3 = left_join(recaps2, distances) %>%
  mutate(SiteType = paste(Type, Site))

recaps3
```

Plot to visually look at mean and variatoin.

```{r}
ggplot(recaps3, aes(x = SiteType, y = Distance))+ geom_boxplot(aes(fill = Type))+
  ylab("Distance from Release Site (m)")+
  theme_bw()
```

Map with color variation by distance, to make sure I got it right.

```{r}
recapsf = st_as_sf(filter(recaps3, !is.na(`Longitude`)), coords = c("Longitude", "Latitude"), crs = 4326)

ggplot()+
  geom_sf(data = WW_Delta)+
  geom_sf(data = recapsf, aes(color = Distance), size =3)+
  geom_sf(data = releasesf, aes(shape = Type), size = 5)+
  scale_color_viridis_c(option = "B")+
  coord_sf(xlim = c(-121.5, -122.2), ylim = c(37.8, 38.5))

recapsum = group_by(recaps3, Site, Type, SiteType) %>%
  summarize(Nsmelt = n(), MeanDist = mean(Distance, na.rm = T), sdDist = sd(Distance, na.rm =T))

recapsum

```

NOw we move on to distance bewteen recaptured fish. I'm doing it oen release at a tiem again.

```{r, eval=FALSE}

RVhdist = Waterdist(Water_map = spacetools::Delta, Points = RVhardt, Latitude_column = `Latitude`,
                    Longitude_column = `Longitude`, PointID_column = fishID)

RVhdistx = Waterdist(Water_map = spacetools::Delta, Points = RVhard, Latitude_column = `Latitude`,
                    Longitude_column = `Longitude`, PointID_column = fishID)

SCsoft  = Waterdist(Water_map = spacetools::Delta, Points = Dwscsoft, Latitude_column = `Latitude`,
                       Longitude_column = `Longitude`, PointID_column = fishID)

SChard = Waterdist(Water_map = spacetools::Delta, Points = Dwschard, Latitude_column = `Latitude`,
                   Longitude_column = `Longitude`, PointID_column = fishID)

RVhdist
RVhdistx
SCsoft
SChard

```

I'm not quite sure the best way to summarize this. I'll start by just the mean of the distance matrix, but i might want do just half the matrix. I'll look into that later.

```{r}

Cluster = data.frame(Release = c("RV Hard", "RV Hard Trailer", "DWSC hard", "DWSC soft"),
                     meadist = c(mean(RVhdistx), mean(RVhdist), mean(SCsoft), mean(SChard)),
                     SD = c(sd(RVhdistx), sd(RVhdist), sd(SCsoft), sd(SChard)))
Cluster = mutate(Cluster, CV = SD/meadist)
Cluster
```