### 2023-03-17 histopathology.Rmd

---
title: "Smelt Cage Histopathology"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readxl)
library(car)
library(emmeans)
library(effects)
library(ordinal)
library(sure)

```

## Stats

```{r cars}
#start with liver lesions
livers = read_excel("histopathology data caging study_to share with dwr.xlsx")

#
ggplot(livers, aes(x = season, y = `liver lesion score`, fill = treatment))+
  geom_boxplot()

#quick linear model
goodlivers = filter(livers, !treatment %in% c("Suisun Marsh", "Yolo Bypass", "DWSC")) %>%
  mutate(treatment = case_when(treatment == "RV" ~ "Rio Vista",
                               TRUE ~ treatment),
         lls = log(`liver lesion score`+1))

m1 = lm(`liver lesion score`~ treatment*season, data = goodlivers)
summary(m1)

```

Interaction not significant, let's drop it.

```{r}
m2 = lm(`liver lesion score`~ treatment + season, data = goodlivers)
Anova(m2)

plot(m2)
hist(residuals(m2))

```
The QQ plot is a little wonky. Let's try a log transformation

```{r}
m2.1 = lm(lls~ treatment + season, data = goodlivers)
Anova(m2.1)
plot(m2.1)

```
Better, still not great

```{r}

#look at partial effects plot
plot(allEffects(m2.1))

#post hocs
#look just at the effects of treatment
treat = emmeans(m2.1, ~treatment)
treat
plot(treat)

#Now the effects of season
seas = emmeans(m2, pairwise ~season)
seas
plot(seas)

```

We could also compare each deployment individually if we wanted to.

```{r}

goodlivers = mutate(livers, deploy = paste(treatment, season), lls = log(`liver lesion score`+1))

m3 = lm(`liver lesion score`~deploy, data = goodlivers)
plot(m3)
#yuck
m3 = lm(lls~deploy, data = goodlivers)
#better. Maybe I should log transform all the liver models

emmeans(m3, pairwise ~ deploy)
plot(emmeans(m3, pairwise ~ deploy), comparison = T)

```

## Now let's look at glycogen

```{r}
glycogen = read_excel("histopathology data caging study_to share with dwr.xlsx", sheet = "liver GD stats") %>%
  rename(glyDep = `liver glycogen depletion`)

#just Rio Vista and FCCL
gly = filter(glycogen, !treatment %in% c("Suisun Marsh", "Yolo Bypass", "DWSC")) %>%
  mutate(treatment = case_when(treatment == "RV" ~ "Rio Vista",
                               TRUE ~ treatment))

#a few quick exploritory plots
ggplot(glycogen, aes(x = treatment, y = glyDep, fill = season)) +
  geom_boxplot()

ggplot(gly, aes(x = treatment, y = glyDep, fill = season)) +
  geom_boxplot()
```

Now the statistics

```{r}
g1 = lm(glyDep~ season + treatment, data = gly)
plot(g1)
```

Not too bad.

```{r}
summary(g1)
Anova(g1)
plot(allEffects(g1))
emmeans(g1, pairwise ~ season)
plot(emmeans(g1, pairwise ~ season), comparison = T)
```

So, winter is definitely lower than fall and summer, FCCL definitely lower than Rio Vista

## Gill lesions

```{r}

gills = read_excel("histopathology data caging study_to share with dwr.xlsx", sheet = "gill lesions stats") %>%
  rename(gill = `gill lesion score`)

goodgills = filter(gills, !treatment %in% c("Suisun Marsh", "Yolo Bypass", "DWSC")) %>%
  mutate(treatment = case_when(treatment == "RV" ~ "Rio Vista",
                               TRUE ~ treatment))

#exploritory plots
ggplot(gills, aes(x = treatment, y = gill, fill = season)) +
  geom_boxplot()

ggplot(goodgills, aes(x = treatment, y = gill, fill = season)) +
  geom_boxplot()
```

Now the statistics

```{r}
gl1 = lm(gill~ season + treatment, data = goodgills)
plot(gl1)
```
Yuck. need some sort of transformation on that

```{r}

#try a log transformation
goodgills = mutate(goodgills, gls = log(gill+1))
gl1 = lm(gls~ season + treatment, data = goodgills)
plot(gl1)
#still yuck

#square root transformation

goodgills = mutate(goodgills, gls = sqrt(gill))
gl1 = lm(gls~ season + treatment, data = goodgills)
plot(gl1)
#still yuck.

```

Maybe an ordinal regression would be more appropriate since it's integers between zero and 4?

```{r}

library(ordinal)

goodgills = mutate(goodgills, Fgill = factor(gill, ordered = T))
c1 = clm(Fgill ~ treatment + season, data = goodgills)
summary(c1)

convergence(c1)
library(sure)
autoplot.clm(c1)
#beautiful!

#so, no real difference in gill leasions between treatments or seasons

```