### 2023-04-05 smelt_2019_allseasons/analyze_growth.Rmd
Reran length and weight models for hybrid manuscript. Saved some figures and tables from the results. Some of the data might have been re-saved accidentally.

---
title: "analyze_growth"
author: "Catarina Pien"
date: '2022-09-16'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(readr)
library(here)
library(lme4)
library(car)
root <- "smelt_2019_allseasons"
```

# Condition Factor, Length, Weight
1. Read data

* a, e = wrap
* b, f = large. F = lots of dead smelt
* c, d = small

* We can keep c, d, b
```{r}
data <- read_csv(here(root,"data_clean","clean_smelt_cf_2019.csv")) %>%
  filter(!(grepl("FCCL", Cage) & Season == "Winter"),
         !(Cage %in% c("A", "E","F") & Season == "Winter")) %>%
  filter(!(Site == "RV" & is.na(Cage))) %>%
mutate(Season = factor(Season)) %>%
  mutate(CageRep = paste0(Season, "_", Site, "_", Cage))

# Note there are a few fish in the summer that are cage-less. They may need to be identified as to which cage they should belong to??

str(data)
RV = filter(data, Site == "RV")
DWSC = filter(data, Site == "DWSC")
```

Weight pre-posts by season
```{r}
Summer <- data %>% filter(Season == "Summer") %>%
  mutate(Period = factor(Period, levels = c("Pre", "Post")))
Summer_wide <- Summer %>%
  pivot_wider(names_from = "Period",
              values_from = c("FL_cm", "Weight_g", "K"))

Fall <- data %>% filter(Season == "Fall")%>%
  mutate(Period = factor(Period, levels = c("Pre", "Post")))

Fall_wide <- Fall %>%
  pivot_wider(names_from = "Period",
              values_from = c("FL_cm", "Weight_g", "K"))
```

```{r}
hist(Summer$FL_cm_Pre)
hist(filter(Summer, Site=="RV")$FL_cm_Post)
hist(Summer$Weight_g_Pre)
hist(filter(Summer, Site=="RV")$Weight_g_Post)

hist(filter(data, Site == "RV")$FL_cm)
hist(filter(data, Site == "RV")$Weight_g)
hist(filter(data, Site == "RV")$K)

hist(filter(data, Site == "DWSC")$FL_cm)
hist(filter(data, Site == "DWSC")$Weight_g)
hist(filter(data, Site == "DWSC")$K)

hist(Summer$FL_cm)
hist(Summer$Weight_g)
hist(sqrt(Summer$Weight_g))
hist(Summer$K)

hist(Fall$FL_cm)
hist(Fall$Weight_g)
hist(Fall$K)

shapiro.test(filter(data, Site == "DWSC")$FL_cm)
shapiro.test(filter(data, Site == "DWSC")$Weight_g)
shapiro.test(filter(data, Site == "DWSC")$K)

leveneTest(DWSC$FL_cm ~ DWSC$Period, data = data)
leveneTest(DWSC$Weight_g ~ DWSC$Period, data = data)
leveneTest(DWSC$K ~ DWSC$Period, data = data)

shapiro.test(Summer$FL_cm)
shapiro.test(sqrt(Summer$Weight_g))
shapiro.test(Summer$K)

leveneTest(Summer$FL_cm~Summer$Period)
leveneTest(Summer$Weight_g~Summer$Period)
leveneTest(Summer$K~Summer$Period)

shapiro.test(Fall$FL_cm)
shapiro.test(Fall$Weight_g)
shapiro.test(Fall$K)

leveneTest(Fall$FL_cm~Fall$Period)
leveneTest(Fall$Weight_g~Fall$Period)
leveneTest(Fall$K~Fall$Period)
```

## Plots
```{r}
Summer$Period = factor(Summer$Period, levels = c("Pre", "Post"))
ggplot(Summer) + geom_boxplot(aes(Site, FL_cm, fill = Period)) 
ggplot(Summer) + geom_boxplot(aes(Site, Weight_g, fill = Period)) 
ggplot(Summer) + geom_boxplot(aes(Site, K, fill = Period))

Fall$Period = factor(Fall$Period, levels = c("Pre", "Post"))
ggplot(Fall) + geom_boxplot(aes(Site, FL_cm, fill = Period))
ggplot(Fall) + geom_boxplot(aes(Site, Weight_g, fill = Period))
ggplot(Fall) + geom_boxplot(aes(Site, K, fill = Period)) 
```

# T-tests

## Summer
Pre-post model (t-test)
```{r}
data_sFCCL <- Summer %>%
  filter(Site == "FCCL")
data_sRV <- Summer %>%
  filter(Site == "RV" | (Site == "FCCL" & Period == "Pre"))

# t-tests ---------------------------------------
t.test(filter(data_sFCCL,Period=="Pre")$Weight_g, filter(data_sFCCL,Period=="Post")$Weight_g)
t.test(filter(data_sFCCL,Period=="Pre")$FL_cm, filter(data_sFCCL,Period=="Post")$FL_cm) 
t.test(filter(data_sFCCL,Period=="Pre")$K, filter(data_sFCCL,Period=="Post")$K)

t.test(filter(data_sRV,Period=="Pre")$Weight_g, filter(data_sRV,Period=="Post")$Weight_g)
t.test(filter(data_sRV,Period=="Pre")$FL_cm, filter(data_sRV,Period=="Post")$FL_cm) 
t.test(filter(data_sRV,Period=="Pre")$K, filter(data_sRV,Period=="Post")$K)

# Linear models: resids look good
# FL -------------------------------------------------
(m_fccl_summer1 <- lm(FL_cm~Period, data = data_sFCCL))
summary(m_fccl_summer1)
anova(m_fccl_summer)
par(mfrow = c(2,3))
plot(m_fccl_summer1)
hist(resid(m_fccl_summer1))
boxplot(FL_cm~Period, data_sFCCL)

(m_rv_summer1 <- lm(FL_cm~Period, data = data_sRV))
par(mfrow = c(2,3))
plot(m_rv_summer1)
hist(resid(m_rv_summer1))
boxplot(FL_cm~Period, data_sRV)

# Weight ----------------------------------------------
(m_fccl_summer2 <- lm(Weight_g~Period, data = data_sFCCL))
summary(m_fccl_summer2)
anova(m_fccl_summer2)
par(mfrow = c(2,3))
plot(m_fccl_summer2)
hist(resid(m_fccl_summer2))
boxplot(Weight_g~Period, data_sFCCL)

(m_rv_summer2 <- lm(Weight_g~Period, data = data_sRV))
summary(m_rv_summer2)
par(mfrow = c(2,3))
plot(m_rv_summer2)
hist(resid(m_rv_summer2))
boxplot(Weight_g~Period, data_sRV)

# K ------------------------------------------------
(m_fccl_summerK <- lm(K~Period, data = data_sFCCL))
summary(m_fccl_summerK)
anova(m_fccl_summerK)
par(mfrow = c(2,3))
plot(m_fccl_summerK)
hist(resid(m_fccl_summerK))
boxplot(K~Period, data_sFCCL)

(m_rv_summerK <- lm(K~Period, data = data_sRV))
summary(m_rv_summerK)
par(mfrow = c(2,3))
plot(m_rv_summerK)
hist(resid(m_rv_summerK))
boxplot(K~Period, data_sRV)

```

## Fall
```{r}
data_fFCCL <- Fall %>%
  filter(Site == "FCCL")
data_fRV <- Fall %>%
  filter(Site == "RV" | (Site == "FCCL" & Period == "Pre"))
data_fSM <- Fall %>%
  filter(Site == "SM" | (Site == "FCCL" & Period == "Pre"))
data_fYolo <- Fall %>%
  filter(Site == "Yolo" | (Site == "FCCL" & Period == "Pre"))

# t-tests ---------------------------------------
t.test(filter(data_fFCCL,Period=="Pre")$Weight_g, filter(data_fFCCL,Period=="Post")$Weight_g)
t.test(filter(data_fFCCL,Period=="Pre")$FL_cm, filter(data_fFCCL,Period=="Post")$FL_cm) 
t.test(filter(data_fFCCL,Period=="Pre")$K, filter(data_fFCCL,Period=="Post")$K)

t.test(filter(data_fRV,Period=="Pre")$Weight_g, filter(data_fRV,Period=="Post")$Weight_g)
t.test(filter(data_fRV,Period=="Pre")$FL_cm, filter(data_fRV,Period=="Post")$FL_cm) 
t.test(filter(data_fRV,Period=="Pre")$K, filter(data_fRV,Period=="Post")$K)

t.test(filter(data_fSM,Period=="Pre")$Weight_g, filter(data_fSM,Period=="Post")$Weight_g)
t.test(filter(data_fSM,Period=="Pre")$FL_cm, filter(data_fSM,Period=="Post")$FL_cm) 
t.test(filter(data_fSM,Period=="Pre")$K, filter(data_fSM,Period=="Post")$K)

t.test(filter(data_fYolo,Period=="Pre")$Weight_g, filter(data_fYolo,Period=="Post")$Weight_g)
t.test(filter(data_fYolo,Period=="Pre")$FL_cm, filter(data_fYolo,Period=="Post")$FL_cm) 
t.test(filter(data_fYolo,Period=="Pre")$K, filter(data_fYolo,Period=="Post")$K)

# Linear models: resids look good
# FL -------------------------------------------------
(m_fccl_fall1 <- lm(FL_cm~Period, data = data_fFCCL))
summary(m_fccl_fall1)
anova(m_fccl_fall)
par(mfrow = c(2,3))
plot(m_fccl_fall1)
hist(resid(m_fccl_fall1))
boxplot(FL_cm~Period, data_fFCCL)

(m_rv_fall1 <- lm(FL_cm~Period, data = data_fRV))
par(mfrow = c(2,3))
plot(m_rv_fall1)
hist(resid(m_rv_fall1))
boxplot(FL_cm~Period, data_fRV)

(m_sm_fall1 <- lm(FL_cm~Period, data = data_fSM))
par(mfrow = c(2,3))
plot(m_sm_fall1)
hist(resid(m_sm_fall1))
boxplot(FL_cm~Period, data_fSM)

(m_Yolo_fall1 <- lm(FL_cm~Period, data = data_fYolo))
par(mfrow = c(2,3))
plot(m_Yolo_fall1)
hist(resid(m_Yolo_fall1))
boxplot(FL_cm~Period, data_fYolo)

# Weight ----------------------------------------------
(m_fccl_fall2 <- lm(Weight_g~Period, data = data_fFCCL))
summary(m_fccl_fall2)
anova(m_fccl_fall2)
par(mfrow = c(2,3))
plot(m_fccl_fall2)
hist(resid(m_fccl_fall2))
boxplot(Weight_g~Period, data_fFCCL)

(m_rv_fall2 <- lm(Weight_g~Period, data = data_fRV))
summary(m_rv_fall2)
par(mfrow = c(2,3))
plot(m_rv_fall2)
hist(resid(m_rv_fall2))
boxplot(Weight_g~Period, data_fRV)

(m_sm_fall2 <- lm(Weight_g~Period, data = data_fSM))
summary(m_sm_fall2)
par(mfrow = c(2,3))
plot(m_sm_fall2)
hist(resid(m_sm_fall2))
boxplot(Weight_g~Period, data_fSM)

(m_yolo_fall2 <- lm(Weight_g~Period, data = data_fYolo))
summary(m_yolo_fall2)
par(mfrow = c(2,3))
plot(m_yolo_fall2)
hist(resid(m_yolo_fall2))
boxplot(Weight_g~Period, data_fYolo)

# K ------------------------------------------------
(m_fccl_fallK <- lm(K~Period, data = data_fFCCL))
summary(m_fccl_fallK)
anova(m_fccl_fallK)
par(mfrow = c(2,3))
plot(m_fccl_fallK)
hist(resid(m_fccl_fallK))
boxplot(K~Period, data_fFCCL)

(m_rv_fallK <- lm(K~Period, data = data_fRV))
summary(m_rv_fallK)
par(mfrow = c(2,3))
plot(m_rv_fallK)
hist(resid(m_rv_fallK))
boxplot(K~Period, data_fRV)

(m_sm_fallK <- lm(K~Period, data = data_fSM))
summary(m_sm_fallK)
par(mfrow = c(2,3))
plot(m_sm_fallK)
hist(resid(m_sm_fallK))
boxplot(K~Period, data_fRV)

(m_yolo_fallK <- lm(K~Period, data = data_fYolo))
summary(m_yolo_fallK)
par(mfrow = c(2,3))
plot(m_yolo_fallK)
hist(resid(m_yolo_fallK))
boxplot(K~Period, data_fRV)
```

## Old code
```{r}
t_sRV <- t.test(Summer_wide$FL_cm_Pre, filter(Summer_wide, Site=="RV")$FL_cm_Post) #no
t_sFC <- t.test(Summer_wide$FL_cm_Pre, filter(Summer_wide, Site=="FCCL")$FL_cm_Post) #yes

t.test(Summer_wide$Weight_g_Pre, filter(Summer_wide, Site=="RV")$Weight_g_Post) #on the border
t.test(Summer_wide$Weight_g_Pre, filter(Summer_wide, Site=="FCCL")$Weight_g_Post) #on the border

t.test(Summer_wide$K_Pre, filter(Summer_wide, Site=="RV")$K_Post) #yes
t.test(Summer_wide$K_Pre, filter(Summer_wide, Site=="FCCL")$K_Post) #yes

wilcox.test(Summer_wide$FL_cm_Pre, filter(Summer_wide, Site=="RV")$FL_cm_Post) #no
wilcox.test(Summer_wide$FL_cm_Pre, filter(Summer_wide, Site=="FCCL")$FL_cm_Post) #yes

wilcox.test(Summer_wide$Weight_g_Pre, filter(Summer_wide, Site=="RV")$Weight_g_Post) #no
wilcox.test(Summer_wide$Weight_g_Pre, filter(Summer_wide, Site=="FCCL")$Weight_g_Post) #no

wilcox.test(Summer_wide$K_Pre, filter(Summer_wide, Site=="RV")$K_Post) #yes
wilcox.test(Summer_wide$K_Pre, filter(Summer_wide, Site=="FCCL")$K_Post) #yes
```

```{r}
wilcox.test(Fall_wide$FL_cm_Pre, filter(Fall_wide, Site=="RV")$FL_cm_Post) #no
wilcox.test(Fall_wide$FL_cm_Pre, filter(Fall_wide, Site=="Yolo")$FL_cm_Post) #yes
wilcox.test(Fall_wide$FL_cm_Pre, filter(Fall_wide, Site=="SM")$FL_cm_Post) #yes
wilcox.test(Fall_wide$FL_cm_Pre, filter(Fall_wide, Site=="FCCL")$FL_cm_Post) #yes

wilcox.test(Fall_wide$Weight_g_Pre, filter(Fall_wide, Site=="RV")$Weight_g_Post) #yes
wilcox.test(Fall_wide$Weight_g_Pre, filter(Fall_wide, Site=="Yolo")$Weight_g_Post) #yes
wilcox.test(Fall_wide$Weight_g_Pre, filter(Fall_wide, Site=="SM")$Weight_g_Post) #yes
wilcox.test(Fall_wide$Weight_g_Pre, filter(Fall_wide, Site=="FCCL")$Weight_g_Post) #yes

wilcox.test(Fall_wide$K_Pre, filter(Fall_wide, Site=="RV")$K_Post) #yes
wilcox.test(Fall_wide$K_Pre, filter(Fall_wide, Site=="Yolo")$K_Post) #yes
wilcox.test(Fall_wide$K_Pre, filter(Fall_wide, Site=="SM")$K_Post) #yes
wilcox.test(Fall_wide$K_Pre, filter(Fall_wide, Site=="FCCL")$K_Post) #yes
```

# Models
```{r}
RV_analysis <- data %>%
  group_by(Season, Site, Period, CageRep) %>%
  summarize(meanK = mean(K, na.rm = TRUE),
            meanWt = mean(Weight_g, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(!Site %in% c("SM", "Yolo", "DWSC"))

Post <- filter(RV_analysis, Period == "Post")%>%
  pivot_wider(names_from = "Period", values_from = "meanK")

# Use this to figure out means for batch pre for summer and fall
Pre <- filter(RV_analysis, Period == "Pre") %>%
  pivot_wider(names_from = "Period", values_from = c("meanK", "meanWt"))

# Now bring those values in here
RV_allfish <- data %>%
  filter(!Site %in% c("SM", "Yolo", "DWSC")) %>%
  pivot_wider(names_from = "Period",
              values_from = c("FL_cm", "Weight_g", "K")) %>%
  mutate(K_Pre = case_when(Season == "Fall"~ 0.7214667,
                         Season == "Summer" ~ 0.6626667,
                         Season == "Winter" ~ K_Pre),
         Weight_g_Pre = case_when(Season == "Fall" ~ 1.7213333,
                                Season == "Summer" ~ 0.8206667,
                                Season == "Winter" ~ Weight_g_Pre)) %>%
  mutate(deltaK = (K_Post - K_Pre)/K_Pre,
         deltaWt = (Weight_g_Post - Weight_g_Pre)/Weight_g_Pre) %>%
  filter(Site != "FCCL") %>%
  mutate(Season = factor(Season, levels = c("Winter", "Summer", "Fall"))) %>%
  filter(!is.na(deltaK))

RV_sumfall <- filter(RV_allfish, Season != "Winter") 
RV_sumfall$Season = factor(RV_sumfall$Season, levels = c("Summer", "Fall"))
```

Are the scale of fish weights comparable?
```{r}
ggplot(RV_allfish) + geom_boxplot(aes(Season, Weight_g_Post))
ggplot(RV_allfish) + geom_boxplot(aes(Season, Weight_g_Pre))
```

### Effect of Season at RV
```{r}
library(lmerTest)

# CF
m1 <- lmer(deltaK ~ Season + (1|CageRep), data = RV_allfish)
summary(m1)
anova(m1,type =2)

m1b <- lmer(deltaK ~ Season + (1|CageRep), data = RV_sumfall)
summary(m1b)
plot(m1b)

# Weight
m2 <- lmer(deltaWt~ Season + (1|CageRep), data = RV_allfish)
summary(m2)
plot(m2)

m2b <- lmer(deltaWt ~ Season + (1|CageRep), data = RV_sumfall)
summary(m2b)
plot(m2b) 
```

#### Run post-hoc tests
```{r}
library(emmeans)
m1.emm <- emmeans(m1, "Season", data=RV_allfish)
pairs(m1.emm, adjust = "tukey")
plot(m1.emm, comparisons = TRUE) + theme_bw()

m2.emm <- emmeans(m2, "Season", data=RV_allfish)
pairs(m2.emm, adjust = "tukey")
plot(m2.emm, comparisons = TRUE) + theme_bw()
```

#### Model validation
```{r}
resid.m1 <- resid(m1, type = "pearson")
par(mfrow = c(2,2))
plot(m1)
plot(resid.m1)
hist(resid.m1)
qqnorm(resid.m1)
plot(RV_allfish$Season, resid.m1)

plot(m2)
resid.m2 <- resid(m2, type = "pearson")
par(mfrow = c(2,2))
plot(resid.m2)
hist(resid.m2)
qqnorm(resid.m2)
plot(RV_allfish$Season, resid.m2)

```

#### Plot CI and RE
##### m1
```{r}
library(multcomp)
m1.emm <- emmeans(m1, "Season")
plotm1 = plot(m1.emm, comparisons = TRUE, plotit = F, adjust = "bonferroni")
letter_m1 = cld(m1.emm, Letters = LETTERS,  p.adj.methods = "bonferroni") %>%
  left_join(plotm1) %>%
  mutate(.group = gsub(" ", "", .group))

(plot_season_k <- ggplot(data = letter_m1,) + 
   geom_errorbar(aes(x = Season, ymin = lower.CL, ymax = upper.CL), size = 0.5,  width = 0.1) +

  geom_segment(aes(x = Season, y = emmean, xend = Season, yend = rcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
    geom_segment(aes(x = Season, y = emmean, xend = Season, yend = lcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
     geom_text( aes(label = .group, x = Season, y = upper.CL + 0.01), position = position_dodge2(width = 0.3)) + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_point( aes(x=Season, y=emmean), size = 2.5, shape = 18) +     
  viridis::scale_color_viridis(discrete = TRUE, option = "turbo") +
  labs(y = "Proportional Change in K (g cm^-2) with 95% Confidence Intervals") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        legend.position = "top",
        axis.title.x = element_blank()) )
```

##### m2
```{r}
library(multcomp)
m2.emm <- emmeans(m2, "Season")
plotm2 = plot(m2.emm, comparisons = TRUE, plotit = F, adjust = "bonferroni")
letter_m2 = cld(m2.emm, Letters = LETTERS,  p.adj.methods = "bonferroni") %>%
  left_join(plotm2) %>%
  mutate(.group = gsub(" ", "", .group))

(plot_season_wt <- ggplot(data = letter_m2,) + 
   geom_errorbar(aes(x = Season, ymin = lower.CL, ymax = upper.CL), size = 0.5,  width = 0.1) +

  geom_segment(aes(x = Season, y = emmean, xend = Season, yend = rcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
    geom_segment(aes(x = Season, y = emmean, xend = Season, yend = lcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
  geom_point( aes(x=Season, y=emmean), size = 2.5, shape = 18) +     
   geom_text( aes(label = .group, x = Season, y = upper.CL + 0.02), position = position_dodge2(width = 0.3)) + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
    viridis::scale_color_viridis(discrete = TRUE, option = "turbo") +
  labs(y = "Proportional Change in Weight (g) with 95% Confidence Intervals") +
  theme_bw() +
  theme(axis.text = element_text(size = 11),
        legend.position = "top",
        axis.title.x = element_blank()) )
```

#### Export data and figures
```{r}
# emmeans results -------
write_csv(letter_m1, "smelt_2019_allseasons/data_clean/emmeans_RV_m1.csv")
write_csv(letter_m2, "smelt_2019_allseasons/data_clean/emmeans_RV_m2.csv")

# plots ----------
library(patchwork)
png(filename = "smelt_2019_allseasons/figures/RV_weight_k_season.png", height = 5, width = 9,units = "in", 
    res = 300, pointsize = 12, family = "sans")
plot_season_wt & plot_season_k
dev.off()
```

```{r}
t_ci_m1 <- as.data.frame(confint(m1))[3:5,]
est_winter = 0.034457
ci25 = t_ci_m1[1,1]
ci975 = t_ci_m1[1,2]
ci_m1 <- data.frame(estimate = c(est_winter, est_winter-0.0768, est_winter-0.1844),
                    ci_2_5 = c(ci25, ci25+t_ci_m1[2,1], ci25+t_ci_m1[3,1]),
                    ci_97_5 = c(ci975, ci975+t_ci_m1[2,2], ci975 + t_ci_m1[3,2]))%>%
  mutate(season = c("Winter", "Summer", "Fall"),
         season = factor(season, levels = c("Winter", "Summer", "Fall")))

summary(m1)

ggplot(ci_m1) + 
   geom_errorbar(aes(y = season, xmin = ci_2_5, xmax = ci_97_5), size = 2, color = "aquamarine3", width = 0.1) + 
  geom_point(aes(estimate, season), size = 2.5) +

  labs(x = "Delta K with 95% Confididence Intervals") +
  theme_bw() +
  theme(axis.title.y = element_blank())

re_m1 <- as.data.frame(ranef(m1))

ggplot(re_m1) + 
  geom_point(aes(grp, condval)) +
  geom_errorbar(aes(x = grp, ymin = condval-condsd, ymax = condval+condsd), width = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95),
        axis.title = element_blank())
```

##### m2
```{r}
t_ci_m2 <- as.data.frame(confint(m2))[3:5,]
est_winter = 0.20763
ci25 = t_ci_m2[1,1]
ci975 = t_ci_m2[1,2]
ci_m2 <- data.frame(estimate = c(est_winter, est_winter-0.27075, est_winter-0.60332),
                    ci_2_5 = c(ci25, ci25+t_ci_m2[2,1], ci25+t_ci_m2[3,1]),
                    ci_97_5 = c(ci975, ci975+t_ci_m2[2,2], ci975 + t_ci_m2[3,2]))%>%
  mutate(season = c("Winter", "Summer", "Fall"),
         season = factor(season, levels = c("Winter", "Summer", "Fall")))

summary(m2)

ggplot(ci_m2) + 
   geom_errorbar(aes(y = season, xmin = ci_2_5, xmax = ci_97_5), size = 2, color = "aquamarine3", width = 0.1) + 
  geom_point(aes(estimate, season), size = 2.5) +

  labs(x = "Delta Weight with 95% Confididence Intervals") +
  theme_bw() +
  theme(axis.title.y = element_blank())

re_m2 <- as.data.frame(ranef(m2))

ggplot(re_m2) + 
  geom_point(aes(grp, condval)) +
  geom_errorbar(aes(x = grp, ymin = condval-condsd, ymax = condval+condsd), width = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95),
        axis.title = element_blank())
```

### Effect of Site in the Fall 
```{r}
Fallfish_avg <- data %>%
  group_by(Season, Site, Period, CageRep) %>%
  summarize(meanK = mean(K, na.rm = TRUE),
            meanWt = mean(Weight_g, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(Season == "Fall")

# Now bring those values in here
Fallfish <- data %>%
  filter(Season == "Fall") %>%
  pivot_wider(names_from = "Period",
              values_from = c("FL_cm", "Weight_g", "K")) %>%
  mutate(K_Pre = 0.7214667,
         Weight_g_Pre = 1.721333,
         deltaK = K_Post - K_Pre,
         deltaWt = Weight_g_Post - Weight_g_Pre) %>%
  filter(!is.na(K_Post))%>%
  mutate(Site = factor(Site))

```

```{r}
library(lmerTest)

# CF ------
m3a <- lm(deltaK ~ Site, data = Fallfish)
summary(m3a)

m3 <- lmer(deltaK ~ Site + (1|CageRep), data = Fallfish, REML = TRUE)
summary(m3)

# Not much difference in residual error by including cage effect. (0.09875 to 0.09787)

# Weight -----
m4a <- lm(deltaWt~ Site, data = Fallfish)
summary(m4a)

m4 <- lmer(deltaWt~ Site + (1|CageRep), data = Fallfish)
summary(m4)

# Not much difference in residual error by including cage effect. (0.471 to 0.466)

```

#### Run post-hoc tests
```{r}
library(emmeans)
m3.emm <- emmeans(m3, "Site", data=Fallfish)
pairs(m3.emm, adjust = "bonferroni")
plot(m3.emm, comparisons = TRUE, type = "response") + theme_bw()

m4.emm <- emmeans(m4, "Site", data=Fallfish)
pairs(m4.emm, adjust = "bonferroni")
plot(m4.emm, comparisons = TRUE) + theme_bw()
```

#### Model validation
```{r}
Fallfish <- Fallfish %>%
  mutate(res3 = resid(m3, type = "pearson"),
         res4 = resid(m4, type = "pearson"))
ggplot(Fallfish) + geom_histogram(aes(res3))
ggplot(Fallfish) + geom_boxplot(aes(Site, res3))

resid.m3 <- as.numeric(resid(m3, type = "pearson"))
plot(m3)
par(mfrow = c(2,2))
plot(resid.m3)
hist(resid.m3)
qqnorm(resid.m3)
boxplot(resid.m3~Fallfish$Site)
summary(resid.m3)

str(Fallfish)

plot(m4)
resid.m4 <- resid(m4, type = "pearson")
par(mfrow = c(2,2))
plot(resid.m4)
hist(resid.m4)
qqnorm(resid.m4)
plot(RV_allfish$Season, resid.m4)
boxplot(resid.m4~Fallfish$Site)

ggplot(Fallfish, aes(factor(Site), res4)) + geom_boxplot() + theme_bw()

```

#### Plot CI and RE

##### m3
```{r}
library(multcomp)
m3.emm <- emmeans(m3, "Site")
plotm3 = plot(m3.emm, comparisons = TRUE, plotit = F, adjust = "bonferroni")
letter_m3 = cld(m3.emm, Letters = LETTERS,  p.adj.methods = "bonferroni") %>%
  left_join(plotm3) %>%
  mutate(.group = gsub(" ", "", .group))

(plot_Site_k <- ggplot(data = letter_m3,) +

  geom_segment(aes(x = Site, y = emmean, xend = Site, yend = rcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
    geom_segment(aes(x = Site, y = emmean, xend = Site, yend = lcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
      geom_errorbar(aes(x = Site, ymin = lower.CL, ymax = upper.CL), size = 0.5,  width = 0.1) +
    geom_text( aes(label = .group, x = Site, y = upper.CL + 0.01), position = position_dodge2(width = 0.3)) + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_point( aes(x=Site, y=emmean), size = 2.5, shape = 18) +     
  viridis::scale_color_viridis(discrete = TRUE, option = "turbo") +
  labs(y = "Change in K (g cm^-2) with 95% Confidence Intervals") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        legend.position = "top",
        axis.title.x = element_blank()) )
```

##### m4
```{r}
m4.emm <- emmeans(m4, "Site")
plotm4 = plot(m4.emm, comparisons = TRUE, plotit = F, adjust = "bonferroni")
letter_m4 = cld(m4.emm, Letters = LETTERS,  p.adj.methods = "bonferroni") %>%
  left_join(plotm4) %>%
  mutate(.group = gsub(" ", "", .group))

(plot_Site_wt <- ggplot(data = letter_m4,) +

  geom_segment(aes(x = Site, y = emmean, xend = Site, yend = rcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
    geom_segment(aes(x = Site, y = emmean, xend = Site, yend = lcmpl, color = .group), size = 6, color = "cadetblue3", lineend = "butt") +
   geom_errorbar(aes(x = Site, ymin = lower.CL, ymax = upper.CL), size = 0.5,  width = 0.1) +
    geom_point( aes(x=Site, y=emmean), size = 2.5, shape = 18) +     
   geom_text( aes(label = .group, x = Site, y = upper.CL + 0.03), position = position_dodge2(width = 0.3)) + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
    viridis::scale_color_viridis(discrete = TRUE, option = "turbo") +
  labs(y = "Change in Weight (g) with 95% Confidence Intervals") +
  theme_bw() +
  theme(axis.text = element_text(size = 11),
        legend.position = "top",
        axis.title.x = element_blank()) )
```

#### Export data and figures
```{r}
# emmeans results -------
write_csv(letter_m3, "smelt_2019_allseasons/data_clean/emmeans_fall_m3.csv")
write_csv(letter_m4, "smelt_2019_allseasons/data_clean/emmeans_fall_m4.csv")

# plots ----------
library(patchwork)
png(filename = "smelt_2019_allseasons/figures/model_fall_site.png", height = 5, width = 9,units = "in", 
    res = 300, pointsize = 12, family = "sans")
plot_Site_wt & plot_Site_k
dev.off()
```

```{r}
t_ci_m3b <- as.data.frame(confint(m3))[3:6,]
est_fccl = -0.06439
ci25 = t_ci_m3[1,1]
ci975 = t_ci_m3[1,2]
ci_m3 <- data.frame(estimate = c(est_fccl, est_fccl-0.08557, est_fccl-0.10488, est_fccl-0.06496),
                    ci_2_5 = c(ci25, ci25+t_ci_m3[2,1], ci25+t_ci_m3[3,1], ci25 + t_ci_m3[4,1]),
                    ci_97_5 = c(ci975, ci975+t_ci_m3[2,2], ci975 + t_ci_m3[3,2],  ci975+t_ci_m3[4,2]))%>%
  mutate(site = c("FCCL", "RV", "SM", "Yolo"),
         site = factor(site, levels = c("FCCL", "RV", "SM", "Yolo")))

summary(m3)

ggplot(ci_m3) + 
   geom_errorbar(aes(y = site, xmin = ci_2_5, xmax = ci_97_5), size = 2, color = "aquamarine3", width = 0.1) + 
  geom_point(aes(estimate, site), size = 2.5) +

  labs(x = "Delta K with 95% Confididence Intervals") +
  theme_bw() +
  theme(axis.title.y = element_blank())

re_m3 <- as.data.frame(ranef(m3))

ggplot(re_m3) + 
  geom_point(aes(grp, condval)) +
  geom_errorbar(aes(x = grp, ymin = condval-condsd, ymax = condval+condsd), width = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95),
        axis.title = element_blank())
```

```{r}
t_ci_m4 <- as.data.frame(confint(m4))[3:6,]
est_fccl = -0.35144
ci25 = t_ci_m4[1,1]
ci975 = t_ci_m4[1,2]
ci_m4 <- data.frame(estimate = c(est_fccl, est_fccl-0.04438, est_fccl-0.19573, est_fccl-0.12013),
                    ci_2_5 = c(ci25, ci25+t_ci_m4[2,1], ci25+t_ci_m4[3,1], ci25 + t_ci_m4[4,1]),
                    ci_97_5 = c(ci975, ci975+t_ci_m4[2,2], ci975 + t_ci_m4[3,2],  ci975+t_ci_m4[4,2]))%>%
  mutate(site = c("FCCL", "RV", "SM", "Yolo"),
         site = factor(site, levels = c("FCCL", "RV", "SM", "Yolo")))

summary(m4)

ggplot(ci_m4) + 
   geom_errorbar(aes(y = site, xmin = ci_2_5, xmax = ci_97_5), size = 2, color = "aquamarine3", width = 0.1) + 
  geom_point(aes(estimate, site), size = 2.5) +

  labs(x = "Delta Weight with 95% Confididence Intervals") +
  theme_bw() +
  theme(axis.title.y = element_blank())

re_m4 <- as.data.frame(ranef(m4))

ggplot(re_m4) + 
  geom_point(aes(grp, condval)) +
  geom_errorbar(aes(x = grp, ymin = condval-condsd, ymax = condval+condsd), width = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95),
        axis.title = element_blank())
```

Attempted some code from Jereme Gaeta's lesson but didn't get this to work.

```{r}
m4
fixef(m3)
coef(summary(m3))
confint(m3, method = "Wald")
library(sjPlot)
plot_model(m3)
tab_model(m3,digits = 3)
```

```{r}
# set up to for loop through the cages
plot_cage = unique(Fallfish$Cage)
col_mat = matrix(data = c(140,81,10, 1,102,94, 118,42,131),
nrow = 3, ncol = 3, byrow = TRUE)
# create a vector that spans the observed range of temperatures for simulations
deltaK_seq = seq(min(Fallfish$deltaK), max(Fallfish$deltaK),
length.out=100)
# Use the effects package to calculate a 95% CI around the grand mean model
library(effects)
m3_eff=Effect(focal.predictors = "Site", mod=m3,
xlevels=list(deltaK=deltaK_seq))

# Use the random effect coefficients to generate random effect predictions
# across the observed range of temperatures
y_F1 = coef(m3)$Cage["B3",1] + coef(m3)$Cage["B3",2]*deltaK_seq
y_F2 = coef(m3)$Cage["B4",1] + coef(m3)$Cage["B4",2]*deltaK_seq
y_F3 = coef(m3)$Cage["B5",1] + coef(m3)$Cage["B5",2]*deltaK_seq

y_A = coef(m3)$Cage["A",1] + coef(m3)$Cage["A",2]*deltaK_seq
y_B = coef(m3)$Cage["B",1] + coef(m3)$Cage["B",2]*deltaK_seq
y_C = coef(m3)$Cage["C",1] + coef(m3)$Cage["C",2]*deltaK_seq

# plot it
par(mfrow=c(1,1), mar=c(4, 4.5, 1, 1)+0.1, oma=rep(0,4))
plot(Fallfish$deltaK ~ Fallfish$Site, type="n", 
     las=1,
ylab="deltaK", xlab = "Site")

for(i in 1:length(plot_cage)){
sub = subset(Fallfish, Fallfish$Cage == plot_cage[i])
points(sub$deltaK ~ sub$Site, pch=20,
col=rgb(col_mat[i,1],col_mat[i,2],col_mat[i,3],
maxColorValue = 255, alpha = 200))
}
lines(deltaK_seq, y_F1, type="l", lwd=1.5,
col=rgb(col_mat[1,1],col_mat[1,2],col_mat[1,3],
maxColorValue = 255, alpha = 225))

lines(deltaK_seq, y_F2, type="l", lwd=1.5,
col=rgb(col_mat[2,1],col_mat[2,2],col_mat[2,3],
maxColorValue = 255, alpha = 225))

lines(deltaK_seq, y_F3, type="l", lwd=1.5,
col=rgb(col_mat[3,1],col_mat[3,2],col_mat[3,3],
maxColorValue = 255, alpha = 225))

lines(deltaK_seq, y_A, type="l", lwd=1.5,
col=rgb(col_mat[3,1],col_mat[3,2],col_mat[3,3],
maxColorValue = 255, alpha = 225))

lines(deltaK_seq, y_B, type="l", lwd=1.5,
col=rgb(col_mat[3,1],col_mat[3,2],col_mat[3,3],
maxColorValue = 255, alpha = 225))

lines(deltaK_seq, y_C, type="l", lwd=1.5,
col=rgb(col_mat[3,1],col_mat[3,2],col_mat[3,3],
maxColorValue = 255, alpha = 225))

polygon(x = c(deltaK_seq, deltaK_seq[100], rev(deltaK_seq), deltaK_seq[1]),
y=c(m3_eff$lower, m3_eff$upper[100],
rev(m3_eff$upper), m3_eff$lower[1]),
col=gray(0.25,0.25),
border=NA)
lines(m3_eff$fit~deltaK_seq, lwd=3.5, col="black")
box(which="outer")
```