### 2023-06-09 smelt_2019_allseasons/Zoop 2019.Rmd
catching up on my git hub

---
librarytitle: "Zoop Plotting"
output: html_document
editor_options: 
  chunk_output_type: console
---
# load packages
```{r}
library(tidyverse)
library(lubridate)
library(Polychrome)
library(zooper)
library(broom)
library(vegan)
```

# load & prep summer/fall data
```{r}

zoop <- read_csv("smelt_2019_allseasons/data_raw/ZoopData_SummerFall_2019.csv")
#zoop <- read_csv("data_raw/ZoopData_SummerFall_2019.csv")

zoop$Date <- mdy(zoop$Date)
zoop$Day <- day(zoop$Date)
zoop$Month <- month(zoop$Date)
zoop$Month <- month.abb[zoop$Month]
zoop$MD <- paste(zoop$Month, zoop$Day, sep="-")

#add ring size (0.5m for 50um net)
zoop$RingSize_m <- 0.5

#add 0s for count with NA
zoop$Count[is.na(zoop$Count)] <- 0

#bring in tow data
aug <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
oct <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")
#aug <- read_csv("data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
#oct <- read_csv("data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")

aug$Date <- mdy(aug$Date)
oct$Date <- mdy(oct$Date)

#condense to columns of interest
aug2 <- aug[c(1,4,22,24:27)]
oct2 <- oct[c(1,4,26,28:31)]

tow <- rbind(aug2,oct2)
tow2 <- tow[!is.na(tow$Location),]

#merge
#rename column
zoop <- zoop %>% rename(Location = "Station")

#make locations match
unique(zoop$Location)
unique(tow2$Location)

zoop[zoop$Location %in% c("STTD", "STTD-YOLO", "YOLO"),]$Location <- "YB"
tow2[tow2$Location %in% c("Yolo"),]$Location <- "YB"

all <- merge(zoop, tow2, by = c('Location', 'Date'))

#calculate abundance -----------------------
all$distance <- (all$Revs)*57560/999999 #57,560 = low speed rotor constant
all$volumesampled <- ((pi*(all$RingSize_m^2))/4)*all$distance

all$samplecount <- (all$TotalVolume_ml/all$Subsampled_ml)*all$Count

#calculate ind/m3
all$ind_m3 <- all$samplecount/all$volumesampled

#remove 0's
caught <- all[!(all$ind_m3==0),]
biomass = read_csv("zooptaxa.csv")
caught = left_join(caught, biomass) %>%
  mutate(bio_m3 = Mass*ind_m3)

#add taxa

caught[caught$ID %in% c("Rotifers", "Copepod nauplii"),]$Taxa <- "Microzoo"

caught[caught$ID %in% c("Ostracods"),]$Taxa <- "Ostracod"

caught[caught$ID %in% c("Bosmina", "Ceriodaphnia", "Diaphanosoma", "Alona", "Chydorus", "Eurycercus", "Monospilus", "Alonella", "Daphnia", "Ilyocryptus", "Camptocercus", "Simocephalus", "Pleuroxus", "Graptoleberis", "Macrothrix"),]$Taxa <- "Cladocera"

caught[caught$ID %in% c("Cyclopoid copepodid", "Tropocyclops prasinus", "Limnoithona tetraspina", "Diacyclops thomasi", "Eucyclops sp.", "Mesocyclops edax", "Microcyclops rubellus", "Tropocylcops prasinus", "Acanthocyclops vernalis adult", "Eucyclops pectinifer", "Paracyclops spp."),]$Taxa <- "Cyclopoid"

caught[caught$ID %in% c("Calanoid copepodid", "Pseudodiaptomus forbesi adult", "Pseudodiaptomus forbesi copepodid", "Eurytemora affinis adult", "Sinocalanus doerrii adult"),]$Taxa <- "Calanoid"

caught[caught$ID %in% c("Harpacticoids"),]$Taxa <- "Harpacticoid"

caught[caught$ID %in% c("Maeotias marginata (large)"),]$Taxa <- "Other"
```

# plot summer/fall data
```{r}
#order factors
unique(caught$MD)
caught$MD <- factor(caught$MD, levels=c("Jul-31", "Aug-1", "Aug-2",  "Aug-5", "Aug-6", "Aug-8",  "Aug-12", "Aug-15", "Aug-19", "Aug-22", "Aug-26", "Oct-10", "Oct-11", "Oct-15", "Oct-16", "Oct-18", "Oct-21", "Oct-25", "Oct-28", "Nov-4"))

legend_title <- "Species"

pal36 <- palette36.colors(36)
pal36 <- as.vector(t(pal36))

#First to CPUE
ggplot() +
  geom_bar(position = "fill", stat = "identity", data=caught, aes (x=MD, y=ind_m3, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of all Zoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

#now try BPUE
ggplot() +
  geom_bar(position = "fill", stat = "identity", data=caught, aes (x=MD, y=bio_m3, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Biomass of all Zoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))

#plot with microzoop removed
macro <- caught[!(caught$ID=="Copepod nauplii"),]
macro <- macro[!(macro$ID=="Rotifers"),]

ggplot() +
  geom_bar(position = "fill", stat = "identity", data=macro, aes (x=MD, y=samplecount, fill=ID), colour="black", width = .5) +
  facet_wrap(~Location, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=pal36)+
  labs(x = "Date") +
  labs(y = "Relative Abundance of Macrozoop in Sample")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10),
      axis.text.x = element_text(angle = 90))
```

#bring in winter data
```{r}
#prep of this dataset (code and raw data) can be found on GitHub under Smelt Cages > smelt_2019_winterspring

winter <- read_csv("smelt_2019_allseasons/data_clean/zoop_abundance.csv")
#winter <- read_csv("data_clean/zoop_abundance.csv")

winter$Day <- day(winter$Date)
winter$Month <- month(winter$Date)
winter$Month <- month.abb[winter$Month]
winter$MD <- paste(winter$Month, winter$Day, sep="-")

#remove throw samples
winter <- winter[!(winter$Method=="Throw"),]
#change ostracod taxa
winter[winter$Taxon %in% c("Ostracods"),]$Taxa <- "Ostracod"
#merge with catch df
#condense to columns of interest
winter2 <- winter %>%
  rename(ID = Taxon) %>%
  select(Location, Date, Taxa, samplecount, ind_m3, bio_m3, Day, Month, MD) %>%
  group_by(Location, Date, Taxa, Day, Month, MD) %>%
  summarize(ind_m3 = sum(ind_m3), bio_m3 = sum(bio_m3))

caught2 <- caught %>%
  select(Location, Date, Taxa, samplecount, ind_m3, bio_m3, Day, Month, MD) %>%
  group_by(Location, Date, Taxa, Day, Month, MD) %>%
  summarize(ind_m3 = sum(ind_m3), bio_m3 = sum(bio_m3))

#rename column
#winter2 <- winter2 %>% rename(ID = "Taxon")

#make locations match
unique(winter2$Location)
unique(caught2$Location)

all2019 <- rbind(winter2, caught2)

#remove 0's
caught.19 <- all2019[!(all2019$ind_m3==0),]
```

# plot all 2019
```{r}
#create season column
caught.19$Season <- as.character(NA)

caught.19[caught.19$MD %in% c("Jan-24", "Jan-29", "Jan-31", "Feb-6", "Feb-8", "Feb-12", "Feb-15", "Feb-22", "Feb-28", "Mar-1", "Mar-4", "Mar-8", "Mar-13", "Mar-19", "Mar-25"),]$Season <- "Winter"

caught.19[caught.19$MD %in% c("Jul-31", "Aug-1", "Aug-2",  "Aug-5", "Aug-6", "Aug-8",  "Aug-12", "Aug-15", "Aug-19", "Aug-22", "Aug-26"),]$Season <- "Summer"

caught.19[caught.19$MD %in% c("Oct-10", "Oct-11", "Oct-15", "Oct-16", "Oct-18", "Oct-21", "Oct-25", "Oct-28", "Nov-4"),]$Season <- "Fall"

caught.sea <- caught.19 %>% 
  group_by(Season, Location, Taxa) %>% 
  summarize(sum.count= sum(ind_m3))

#order seasons
caught.sea$Season <- factor(caught.sea$Season, levels=c("Winter", "Summer", "Fall"))

#order locations
caught.sea$Location <- factor(caught.sea$Location, levels=c("DWSC", "RV", "YB", "SM"))

ggplot() +
  geom_bar(position = "fill", stat = "identity", data=caught.sea, aes (x=Location, y=sum.count, fill=Taxa), colour="black", width = .75) +
  facet_wrap(~Season, ncol = 1) +
  theme_bw() + 
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  labs(x = "Location") +
  labs(y = "Relative Abundance")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10))

ggsave("smelt_2019_allseasons/figures/allzoop19.png", dpi=250, height=6, width=6, units="in")

#plot with microzoop removed
caught.sea2 <- caught.sea[!(caught.sea$Taxa=="Microzoo"),]

#create custom palette from viridis to match taxa from diet plot
zoopPalette <- c("#0D0887FF", "#5402A3FF", "#8B0AA5FF", "#B93289FF", "#DB5C68FF", "#F48849FF")

ggplot() +
  geom_bar(position = "fill", stat = "identity", data=caught.sea2, aes (x=Location, y=sum.count, fill=Taxa), colour="black", width = .75) +
  facet_wrap(~Season, ncol = 1) +
  theme_bw() + 
  scale_fill_manual(values=zoopPalette)+
  labs(x = "Location") +
  labs(y = "Relative Abundance")+
  theme(legend.position="right")+
      theme(axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 10))

ggsave("smelt_2019_allseasons/figures/macrozoop19.png", dpi=250, height=6, width=6, units="in")
```

# repeated measures ANOVA
```{r}
#combine taxa 
for.mod<-caught.19 %>% 
  group_by(Location, Date, Taxa, Season) %>% 
  summarize("ind.all"=sum(ind_m3))

shapiro.test(for.mod$ind.all) #very significant, no ANOVA
```

## Multivariate statistics

```{r}

source("PlotNMDS.R")

zoops = mutate(caught.19, SampleID = paste(Location, Date)) %>%
  pivot_wider(id_cols = c(SampleID, Location, Date, MD, Season), names_from = Taxa, values_from = ind_m3, values_fn = sum, values_fill = 0) %>%
  mutate(datatype = "Zoops")

#first let's look at Rio Vista across seasons
RVzoops = filter(zoops, Location == "RV")
fallzoops = filter(zoops, Season == "Fall")
#raw abundance matrix
RVzoopmat = as.matrix(RVzoops[,6:12])
fallzoopmat = as.matrix(fallzoops[,6:12])
#relative abundance matrix
RVzoopmat2 = RVzoopmat/rowSums(RVzoopmat)
fallzoopmat2 = fallzoopmat/rowSums(fallzoopmat)

RVa = adonis2(RVzoopmat~Season, data = RVzoops)
RVa
#

falla = adonis2(fallzoopmat2~Location*datatype, data = fallzoops)

##################################This is the one for the paper
RVb = adonis2(RVzoopmat2~Season, data = RVzoops)
RVb
RVresults = tidy(RVb)
write.csv(RVresults, file = "outputs/RVresults.csv")

#Britt and Bruce wanted biomass instead of abundance

zoopsBM = mutate(caught.19, SampleID = paste(Location, Date)) %>%
  filter(Taxa != "Microzoo") %>%
  pivot_wider(id_cols = c(SampleID, Location, Date, MD, Season), names_from = Taxa, values_from = bio_m3, values_fn = sum, values_fill = 0) %>%
  mutate(datatype = "Zoops")

#first let's look at Rio Vista across seasons
RVzoopsBM = filter(zoopsBM, Location == "RV")

#raw abundance matrix
RVzoopmatBM = as.matrix(RVzoopsBM[,6:11])

#relative abundance matrix
RVzoopmat2BM = RVzoopmatBM/rowSums(RVzoopmatBM)

RVa2 = adonis2(RVzoopmatBM~Season, data = RVzoopsBM)
RVa2
#

RVb = adonis2(RVzoopmat2BM~Season, data = RVzoops)
RVb

```
pretty big differences between seasons, R2 of 54% on the absolute abundance, 64% on relative abundance
pretty big differences between seasons, R2 of 65% on the absolute biomass, 95% on relative biomass
THough I may need to update this with some sort of variable to account for day of year. Not sure.

Now I'll look for changes by location

```{r}
Fallzoops = filter(zoops, Season == "Fall")

#raw abundance matrix
Fallzoopmat = as.matrix(Fallzoops[,6:12])

#relative abundance matrix
Fallzoopmat2 = Fallzoopmat/rowSums(Fallzoopmat)

Falla = adonis2(Fallzoopmat~Location, data = Fallzoops)
Falla
#

######################################################
#this is the second permanova for the paper
Fallb = adonis2(Fallzoopmat2~Location, data = Fallzoops)
Fallb
write.csv(tidy(Fallb), "outputs/FallZoops.csv", row.names = F)

#

```

Wow! Even larger differences in relative and absolute abundance by region

I'm going to try an NMDS just for funzies

```{r}

#raw abundance matrix
zoopmat = as.matrix(zoops[,6:12])

#relative abundance matrix
zoopmat2 = zoopmat/rowSums(zoopmat)

N1 = metaMDS(zoopmat2, trymax = 50, try = 20)
zoops = mutate(zoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(N1, zoops, group = "Season")
PlotNMDS(N1, zoops, group = "Location")

RVN1 = metaMDS(RVzoopmat2)
RVzoops = mutate(RVzoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(RVN1, RVzoops, group = "Season")

```

## Diet data comparison

```{r}
diet2 <- read_csv("smelt_2019_allseasons/data_raw/DSM Cage Diets 2019-2020.csv") #new data, not edited at all
diet2 <- read_csv("data_raw/DSM Cage Diets 2019-2020.csv") #new data, not edited at all

diet2$Date <- mdy(diet2$Date)

#format 
#rename columns
diet2 <- diet2 %>% rename(Cage.ID = "Cage ID", Contents.Weight = "Total Contents Weight", Prey.Taxa = "Prey Taxa", LH.Stage = "LH Stage")

#add taxa column (choose taxa based on what is present in the dataframe)
diet2$Taxa <- as.character(NA)

str(diet2)

diet2[diet2$Prey.Taxa %in% c("Bosmina sp.", "Chydorus sp.", "Daphnia sp.", "Ceriodaphnia sp.", "Chydoridae", "Crangonyx sp.", "Scapholeberis sp.", "Simocephalus sp.", "Eurycercus sp.", "Leydigia sp.", "Cladocera", "Sida sp.", "Alona sp.", "Diaphanosoma sp."),]$Taxa <- "Cladocera"

diet2[diet2$Prey.Taxa %in% c("Americorophium sp.", "Americorophium spinicorne", "Eogammarus confervicolus", "Gammaridea", "Hyallela azteca", "Crangonyx sp.", "Corophiidae", "Amphipoda"),]$Taxa <- "Amphipod"

diet2[diet2$Prey.Taxa %in% c("Pseudodiaptomus forbesi", "Calanoida", "Eurytemora affinis", "Diaptomidae", "Sinocalanus doerrii", "Osphranticum labronectum"),]$Taxa <- "Calanoid"

diet2[diet2$Prey.Taxa %in% c("Acanthocyclops sp.", "Macrocyclops sp.", "Diacyclops thomasi", "Eucyclops sp.", "Cyclopoida", "Halicyclops sp.", "Mesocyclops sp.", "Tropocyclops sp.", "Tropocyclops sp."),]$Taxa <- "Cyclopoid"

diet2[diet2$Prey.Taxa %in% c("Chironomidae", "Diptera"),]$Taxa <- "Diptera"

diet2[diet2$Prey.Taxa %in% c("empty", "EMPTY"),]$Taxa <- "empty"

diet2[diet2$Prey.Taxa %in% c("Harpacticoida"),]$Taxa <- "Harpacticoid"

diet2[diet2$Prey.Taxa %in% c("Ostracoda"),]$Taxa <- "Ostracod"

diet2[diet2$Prey.Taxa %in% c("Fish", "Acari", "Insecta", "Bivalvia", "Culicidae", "egg unidentified", "Corixidae", "Hemiptera", "plant seed", "plant matter", "Unidentified", "Mysidacea", "Cumacea"),]$Taxa <- "Other"

new_DF <- diet2[is.na(diet2$Taxa),] #update code above with new prey.taxa that weren't found in previous diets
#there are two "copepoda" entries with no count or location...don't seem that important so I am going to just go ahead and remove them
diet2[diet2$Taxa %in% c("empty"),]$Count <- "0"
diet2 <- diet2[!is.na(diet2$Count),]

#Tag column currently has rows with just the Tag and others with location, date, and tag
#first break those two kinds of columns apart
weird.tag <- diet2[is.na(diet2$Location),]
norm.tag <- diet2[!is.na(diet2$Location),]

#then for break out the Tag ID into Location and tag (remove date)
weird.tag <- weird.tag %>% separate(Tag, c("Location", "X.date", "Tag"))

#remove new date column
weird.tag <- subset(weird.tag, select = -X.date)

#add mesh types for prototype experiment
#create new column for mesh types
Cage_type <- as.data.frame(matrix(0, ncol = 2, nrow = 6))
colnames(Cage_type) <- c("Cage.ID", "Mesh")

Cage_type$Cage.ID <- c("A", "B", "C", "D", "E", "F")
Cage_type$Mesh <- c("small", "large", "wrap", "wrap", "small", "large")

norm.tag2 <- merge(norm.tag, Cage_type, by = "Cage.ID", all.x = TRUE)

#add mesh types (NA) for summer/fall experiements
weird.tag$Mesh <- NA

#merge the two
diet <- rbind(weird.tag, norm.tag2)

#change location titles 
diet = mutate(diet, Location2 = case_when(Location %in% c("RV", "RIVERS") ~ "RV", 
                                       Location == "Sac DWSC" ~ "DWSC",
                                       Location == "SM" ~ "SM",
                                       Location == "YB" ~ "YB"),
              Month = month(Date),
              Season = case_when(Month %in% c(2,3,5) ~ "Winter",
                                 Month == 8 ~ "Summer",
                                 Month == 11 ~ "Fall"),
              Location = Location2,
              datatype = "diet")

#OMG DON"T PUT COMMENTS IN THE COUNT COLUMN
diet$Count[which(diet$Count == "1 maybe 2")] = "1"
diet$Count = as.numeric(diet$Count)

save(diet, file = "Dietdata.Rdata")

#pivot_wider
dietwide = mutate(diet, SampleID = paste(Location, Tag)) %>%
  pivot_wider(id_cols = c(SampleID, Location, Date, Season, datatype), names_from = Taxa, values_from = Count, values_fn = sum, values_fill = 0)

```

Now put the zoop data and diet data together

```{r}
zoopdiet = bind_rows(zoops, dietwide) %>%
  mutate(Amphipod = case_when(is.na(Amphipod) ~0, TRUE ~ Amphipod),
         Diptera = case_when(is.na(Diptera) ~0, TRUE ~ Diptera)) %>%
  select(-empty, -Microzoo)

rowSums(zoopdiet[,c(6:11, 13, 14)])

#remove empty rows
zoopdiet = zoopdiet[which(rowSums(zoopdiet[,c(6:11, 13, 14)]) !=0),] %>%
  mutate(datatype = as.factor(datatype), Season = as.factor(Season), Location = as.factor(Location))

#Convert to reletive abundance
zoopdietmat = as.matrix(zoopdiet[,c(6:11, 13, 14)])
zoopdietmat2 = zoopdietmat/rowSums(zoopdietmat)

#look at just Rio Vista
RVdietzoops = filter(zoopdiet, Location == "RV")

#Now fall
Falldietzoops = filter(zoopdiet, Season == "Fall")

#Convert to reletive abundance
RVzoopdietmat = as.matrix(RVdietzoops[,c(6:11, 13, 14)])
RVzoopdietmat2 = RVzoopdietmat/rowSums(RVzoopdietmat)

Fallzoopdietmat = as.matrix(Falldietzoops[,c(6:11, 13, 14)])
Fallzoopdietmat2 = Fallzoopdietmat/rowSums(Fallzoopdietmat)

##########################This is the analysis for the paper!!!!
RVda = adonis2(RVzoopdietmat2~ datatype*Season, data = RVdietzoops)
RVda
write.csv(tidy(RVda), "outputs/DietZooppermanova.csv")

Fallda = adonis2(Fallzoopdietmat2~ Location*datatype, data = Falldietzoops)

Fallda
write.csv(tidy(Fallda), "outputs/DietZooppermanovaFall.csv")

```

Let's look at this a little more closely,s ee if we see any patterns.

```{r}
#It's really odd that data type isn't that important given some critters that show up in diet and not zooplankton at all.

ggplot(zoopdietlong, aes(x = SampleID, y = count, fill = Taxa)) + geom_col()+
  facet_wrap(~datatype, scales = "free")

ggplot(filter(zoopdietlong, Location == "RV"), aes(x = SampleID, y = count, fill = Taxa)) + geom_col()+
  facet_wrap(Season~datatype, scales = "free")

ggplot(filter(zoopdietlong, Location == "RV"), aes(x = SampleID, y = count, fill = Taxa)) + geom_col(position = "fill")+
  facet_wrap(Season~datatype, scales = "free")

```

I tested for an interaction between datatype and season, but nothing. I think it's just too variable to say much.

```{r}
#NMDS's that we probably won't include

zdM = metaMDS(zoopdietmat2)

PlotNMDS(zdM, data = zoopdiet, group = "datatype")

PlotNMDS(zdM, data = zoopdiet, group = "Location")

RVM = metaMDS(RVzoopdietmat2, trymax = 100)

PlotNMDS(RVM, data = RVdietzoops, group = "Season")
```

Basic plot of zoops versus diet

```{r}
zoopdietlong = pivot_longer(zoopdiet, cols = c(Ostracod, Cyclopoid, Calanoid, Cladocera, Harpacticoid, Other, Amphipod, Diptera),
                            names_to = "Taxa", values_to= "count")

zoopnos = group_by(zoopdiet, datatype, Season, Location) %>%
  summarize(N = length(unique(SampleID)))

ggplot(zoopdietlong, aes(x = datatype, y = count, fill = Taxa))+
  geom_col(position = "fill")+
  facet_grid(Location~Season)+
  scale_fill_brewer(palette = "Dark2", name = "Taxon")+
  theme_bw()+
  ylab("Relative Abundance")+
  geom_label(data = zoopnos, aes(x = datatype, y = 1, label = N), inherit.aes = FALSE)+
  scale_x_discrete(labels = c("Zooplankton", "Diets"), name = NULL)

ggsave("outputs/zoopdiet.tiff", device = "tiff", width = 8, height = 6.5)

```

Can I do a permanova that takes date and location into accoutn?

```{r}

zda = adonis2(zoopdietmat2 ~ datatype*Season + Location, strata = zoopdiet$Location, data = zoopdiet)
zda

zda1 = adonis2(zoopdietmat2 ~ datatype*Season + Location,data = zoopdiet)
zda1

with(zoopdiet, adonis2(zoopdietmat2 ~ datatype+ Season , strata = Location, data = zoopdiet, permutations = 199))

with(zoopdiet, adonis2(zoopdietmat2 ~ datatype+ Season +Location, strata = Season, data = zoopdiet, permutations = 199))

```