### 2023-06-09 smelt_2019_allseasons/clean_conditionfactor_survival.Rmd

---
title: "Clean_Condition_Survival"
author: "Catarina Pien"
date: "12/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
This code takes raw summer, fall, winter data and combines it into one file of length, weight, condition factor. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(readr)
library(here)
root <- "smelt_2019_allseasons"
```

Tasks:
1. Read winter/spring, summer/fall data in
2. Clean names and bind together
3. Calculate Pre and Post condition factors
4. Make survival file

# Condition Factor, Length, Weight
1. Read data
```{r}
summerdata <- read_csv(here(root,"data_raw","Raw_Smelt-Length-Weight_Summer2019.csv"), col_types = cols(Deployment = "f", Period = "f", Site = "f", ID = "c", FL_mm = "c", Weight_g= "d", Cage = "f", K =  "d", Comments =  "c")) %>%
  rename(K0 = K) %>%
  mutate(Season = "Summer")

falldata <- read_csv(here::here(root,"data_raw", "Raw_Smelt-Length-Weight_Fall2019.csv"), col_types = cols(Deployment = "f", Period = "f", Site = "f", ID = "c", FL_mm = "c", Weight_g= "d", Cage = "f", K =  "d", Comments =  "c")) %>%
  rename(K0 = K) %>%
  mutate(Season = "Fall")

winterspringdata <- read_csv(here(root, "data_raw", "Cagetype_data.csv")) %>%
  rename(Pre_FL=Pre_FL_cm,
         Post_FL=Post_FL_cm,
         Pre_Weight=Pre_Weight_g,
         Post_Weight=Post_Weight_g) %>%
  mutate(Season = "Winter",
         ID = paste0("Winter_",`...1`))
```

2. Reformat
```{r}
winspring_long <- winterspringdata %>% 
  select(-Delta_CF, -`...1`) %>%
  pivot_longer(cols = c(Pre_FL:Post_CF),
               names_to = c("Period", ".value"),
               names_sep = "_")

# Filter to only large mesh and RV
winspring_clean <- winspring_long %>%
  rename(FL_cm = FL,
         Weight_g = Weight,
         K0 = CF) %>%
  mutate(Deployment = "Winter",
         Deployment = as.factor(Deployment),
         Period = as.factor(Period),
         Site = as.factor(Site),
         Cage = as.factor(Cage))%>%
  select(Deployment, Period, Site, ID, FL_cm, Weight_g, Cage, Season)

summerfall_clean <- rbind(summerdata, falldata) %>%
  mutate(FL_mm = replace(FL_mm, FL_mm == "-", NA),
         FL_mm = as.numeric(FL_mm),
         FL_mm = FL_mm/10, na.rm = TRUE) %>%
  rename(FL_cm = FL_mm) %>%
  select(Deployment, Period, Site, ID, FL_cm, Weight_g, Cage, Season)

```

Bind 
```{r}
CF_2019 <- rbind(winspring_clean, summerfall_clean) %>%
  mutate(K= round(100 * Weight_g/(FL_cm^3),3)) 
```

Write file
```{r}
write.csv(CF_2019, "smelt_2019_allseasons/data_clean/clean_smelt_cf_2019.csv", row.names = FALSE)
```

# Survival
```{r}
Survival <- data.frame(Deployment = factor(), Site = factor(), Survived = numeric(), Total = numeric())

```

Winter
RV: 60, 64, 64, 64, 62, 56 / 64

Spring
DWSC: 59, 60, 58, 60, 60, 58 / 60

Summer
Yolo: 0 / 60
RV: 152 fish on data sheet - 21 died in cage / 60 60 60

Fall
Yolo: 
RV: 161 fish on data sheet - 5 cage mortality in comments = 156/ 60 60 60
SM: 177 fish on data sheet - 5 cage mortality in comments/ 60 60 60
YB: 161 fish on data sheet - 2 "Dead in cage) = 159/ 60 60 60

Original table: https://cawater.sharepoint.com/sites/dwrdoc-aes/Shared%20Documents/Forms/AllItems.aspx?viewid=5f80dd60%2D5e8b%2D4f46%2Dab49%2D609d1052bd85&id=%2Fsites%2Fdwrdoc%2Daes%2FShared%20Documents%2FDelta%20Smelt%20Enclosure%20Study%2FPermits%2FAnnual%20Reports%2F2019%2FDelta%20Smelt%20Enclosure%20Study%202019%20Annual%20Report%20FINAL%2Epdf&parent=%2Fsites%2Fdwrdoc%2Daes%2FShared%20Documents%2FDelta%20Smelt%20Enclosure%20Study%2FPermits%2FAnnual%20Reports%2F2019

Info for this data: "V:\Data\Summer'19\August Measurements and Dissections" & "V:\Data\Summer'19\October Measurements and Dissections"
```{r}
cage <- read.csv("smelt_2019_allseasons/data_raw/Cagetype_data.csv")
# Cages B + F are large mesh
# Winter
(64+56)/(64+64) # RV 0.94

# Summer
(152-21)/180 #RV 0.73

#Fall
(161-5)/180 #RV 0.87
(177-5)/180 #SM 0.96
(161-2)/180 #YB 0.88
```

Make survival table
```{r}
Cage_survival <- as.data.frame(matrix(0, ncol = 4, nrow = 6))
colnames(Cage_survival) <- c("Deployment", "Site", "Survived", "TotalFish")
Cage_survival$Deployment = c("Winter", "Summer", "Summer", "Fall", "Fall", "Fall")
Cage_survival$Site = c("RV", "RV", "Yolo", "RV", "SM", "Yolo")
Cage_survival$Survived = c(120, 131, 0, 156, 172, 159)
Cage_survival$TotalFish = c(128, 180, 60, 180, 180, 180)

Cage_survival_total <- Cage_survival %>%
  mutate(Prop = round(Survived/TotalFish,3))

```

```{r}
write.csv(Cage_survival_total, "smelt_2019_allseasons/data_clean/cage_survival_table.csv", row.names = FALSE)
```