### 2023-11-22 smelt_2019_allseasons/DS CTM 2019-2023.Rmd
Adding biofouling analyses

---
title: "2019 DS CTM"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r load packages & data}
library(ggplot2)
library(tidyverse)
library(ggpubr)

# setwd("C:/Users/bedavis/OneDrive - California Department of Water Resources/My Documents/Special Studies/Internal/Delta Smelt Cages/Cage Synthesis")

```

Summer Fall CTMs in 2019

```{r T 2019  stats}
#test of SM and RVB in fall
field.ctm <- read.csv("DS CTM 2019.csv")

field.ctm.2 <- field.ctm %>% 
  filter(!Season == "Summer")
t.test(CTM ~ Location, data = field.ctm.2)
# t = -1.9141
# df = 15.33
# p-value = 0.074

#test of RV in summer v fall
field.ctm.3 <- field.ctm %>% 
  filter(!Location == "SM")
t.test(CTM ~ Season, data = field.ctm.3)

#plot

ggplot(data=field.ctm,aes(x=Loc.Seas, y=CTM, fill=Location)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.3) +
    geom_jitter(color="black", size=1, alpha=0.5,height=0,width=0.1) +
    theme_bw() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
  scale_y_continuous(breaks = round(seq(min(23), max(30), by = 1.0),1), name="Critical Thermal Maximum (°C)")+
    xlab("")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"))
```

CTM acclimation regression

```{r test for normality}

data <- read.csv("data/CTM Regression 2019.csv")
#density plot
ggdensity(data$mean.ctm)

#QQ plot
ggqqplot(data$mean.ctm)

#shapiro test
shapiro.test(data$mean.ctm)
```

```{r separate data and create models}

#removed the lab outlier from the dataset
data.2 <- data %>% 
  filter(!ID == "f")

#separate data and create separate models

field <- data.2 %>% 
  filter(loc == "Field")

lab <- data.2 %>% 
  filter(loc == "Lab")

#model for field CTM data
field.mod <- lm(mean.ctm ~ acc.temp, field)
anova(field.mod)
summary(field.mod) 
#y = 0.247x + 22.057
#abline(a = 22.057, b = 0.247) 
#rsquared = 0.884
#adjusted rsquared = 0.7681 = 0.77
#p = 0.22

#model for lab CTM data
lab.mod <- lm(mean.ctm ~ acc.temp, lab)
anova(lab.mod)
summary(lab.mod) 
#y = 0.276x + 23.771 
#abline(a = 23.771, b = 0.276)
#rsquared = 0.9289
#adjusted rsquared = 0.9051 = 0.91
#p = 0.008

#estimated CTM of fish deployed at YB during fall
#22.1 + (0.25*16.5) 
# = 26.225

#estimated CTM of fish deployed at SM during summer
#22.1 + (0.25*23.4)
# = 27.95

```

```{r plot regression 2019}

  ggplot(data.2, aes(x = acc.temp, y = mean.ctm, color = loc))+
  geom_smooth(method = "lm", se = F,fullrange=F)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = mean.ctm - temp.sd, ymax = mean.ctm + temp.sd), width = 0.2)+
    scale_color_manual("Study Location",
                    breaks = c("Field", "Lab"),
                    labels = c("Field", "Lab"),
                    values = c("cyan3", "tomato"))+
  geom_point(x = 17, y = 25.4, size = 2, color = "tomato")+
  geom_errorbar(aes(x = 17, ymin = 25.4 - 1.7, ymax = 25.4 + 1.7), color = "tomato", width = 0.2)+
  scale_y_continuous(breaks = round(seq(min(23), max(30), by = 1.0),1), name="Critical Thermal Maximum (°C)")+
  scale_x_continuous(breaks = round(seq(min(12), max(26), by = 2.0),1), name="Acclimation Temperature (°C)")+
  annotate("text", x = 14, y = 28.5, label = "y = 0.28x + 23.8")+
  annotate("text", x = 13.5, y = 28.15, label = "R^2 == 0.91", parse = T)+
  annotate("text", x = 19, y = 26, label = "y = 0.25x + 22.1")+
  annotate("text", x = 18.45, y = 25.65, label = "R^2 == 0.77", parse = T)+
 theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"))+ theme(legend.position="bottom")

```

```{r singe linear model 2019}
##New single linear model with location as factor.

full.mod <- lm(mean.ctm ~ acc.temp+loc, data.2)
summary(full.mod)
#intercept of field fish 21.74
#slope is 0.26, significant effect of acclimation temperature on CTM (p<0.001), with effect of study location (p<0.001).
#lab intercept is 2.22 higher than field which is 23.96
#Adj R2=0.94
anova(full.mod)

modsum = summary(full.mod)
summary(full.mod)

#plot
 ggplot(data.2, aes(x = acc.temp, y = mean.ctm, color = loc))+
 # geom_smooth(method = "lm", se = F,fullrange=F, size=.5)+
   geom_abline(slope = modsum$coefficients[2,1], intercept = modsum$coefficients[1,1], color = "dodgerblue")+
   geom_abline(slope = modsum$coefficients[2,1], intercept = modsum$coefficients[1,1]+modsum$coefficients[3,1], color = "tomato")+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = mean.ctm - temp.sd, ymax = mean.ctm + temp.sd), width = 0.2)+
    scale_color_manual("Study Location",
                    breaks = c("Field", "Lab"),
                    labels = c("Field", "Lab"),
                    values = c("dodgerblue", "tomato"))+
  geom_point(x = 17, y = 25.4, size = 2.5, shape=1,color = "tomato")+
  geom_errorbar(aes(x = 17, ymin = 25.4 - 1.7, ymax = 25.4 + 1.7), color = "tomato", width = 0.2)+
  geom_point(x = 16.5, y = 26.5, size = 2.2,shape= 15,color = "dodgerblue")+
  geom_errorbar(aes(x = 16.5, ymin = 26.5 - 0.86, ymax = 26.5 + 0.86), color = "dodgerblue", width = 0.2)+
  scale_y_continuous(breaks = round(seq(min(23), max(30), by = 1.0),1), name="Critical Thermal Maximum (°C)")+
  scale_x_continuous(breaks = round(seq(min(12), max(26), by = 2.0),1), name="Acclimation Temperature (°C)")+
  annotate("text", x = 21.5, y = 24, label = "R^2 == 0.94", parse = T)+
   annotate("text", x = 20, y = 27.2, label = "CTmax = 0.26x + 21.73", angle =18)+
   annotate("text", x = 14, y = 27.9, label = "CTmax = 0.26x + 23.95", angle =18)+
  annotate("text", x = 21.5, y = 23.5, label = " P < 0.001", parse=T)+
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"), legend.position=c(0.15,0.85))

 ggsave("plots/CTMs.tiff", device = "tiff", width =6, height =5)

```

Fall 2023 CTMs

```{T  2023}
#call in all CTMs
ctm <- read.csv("ALL CTM.csv")
#filter to just 2023
c1 <- ctm %>% 
  filter(!Season == "2019 Summer")
c2 <- c1 %>% 
  filter(!Season == "2019 Fall")

#plot size of those recovered or not
ggplot(c2, aes(x=FL, y=Mass, color=Recovered, shape=Treatment))+
  geom_point()
ggplot(c2, aes(x=FL, y=CTM, color=Recovered))+
  geom_point()
ggplot(c2, aes(x=Mass, y=CTM, color=Recovered))+
  geom_point()
ggplot(c2, aes(x=Mass, y=CTM, color=Cage))+
  geom_point()

#remove those that didnt survive
c3 <- c2 %>% 
  filter(Recovered == "y")

#plot 2023 CTM data
ggplot(c3, aes(x=Site, y=CTM, color=Season))+
  geom_boxplot()
ggplot(c3, aes(x=Treatment, y=CTM, color=Site))+
  geom_boxplot()
ggplot(c3, aes(x=as.factor(Cage), y=CTM, color=Site))+
  geom_boxplot()
#remove 20C outlier in marsh
c3 <- c3 %>% 
  filter(!Chamber == "P")

#remake fig
ggplot(c3, aes(x=as.factor(Cage), y=CTM, color=Site))+
  geom_boxplot()+ylim(20,30)

ggplot(data=c3,aes(x=Treatment, y=CTM, fill=Site)) +
    geom_boxplot(outlier.shape=NA) +
    scale_fill_viridis(discrete = TRUE, alpha=0.3) +
    scale_color_viridis(discrete = TRUE, alpha=0.3) +

    geom_jitter( position=position_jitterdodge(jitter.width =.1), size=1, alpha=0.5,
                shape = 21) +
    theme_bw() +
  scale_y_continuous(breaks = round(seq(min(23), max(30), by = 1.0),1), name="Critical Thermal Maximum (°C)")+
    xlab("")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"))

#add 2019 data to figure and only scrub data

c1a <- c1 %>% 
  filter(Recovered == "y")
c1b <- c1a %>% 
  filter(!Treatment == "flip")
ggplot(data=c1b,aes(x=Season, y=CTM, fill=Site)) +
    geom_boxplot(outlier.shape=NA) +
    scale_fill_viridis(discrete = TRUE, alpha=0.5,direction=1,option="E") +
    geom_point(position=position_jitter(width = 0.1), color="black",size=1, alpha=0.5, aes(group = Season)) +
  facet_wrap(~Site)+
    xlab("")+
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"))+
  scale_y_continuous(breaks = round(seq(min(20), max(30), by = 2.0),1), name="Critical Thermal Maximum (°C)")

#run model of 2023 without control
c3a <- c3 %>% 
  filter(!Treatment == "control")

#normality test
model1=lm(CTM ~ Site*Treatment+(1|Cage), data=c3a)
Anova(model1)

library(emmeans)
pairs(emmeans(model1, ~Site:Treatment))

model2=lm(CTM ~ Site+Treatment, data=c3a)
Anova(model2)
plot(model2)
hist(c3a$CTM)

#get 2023 mean values for later regression table
ctm2 <- ctm %>% 
  filter(Recovered == "y")
ctm2 <- ctm %>% 
  filter(!Chamber == "P")

ctm.sum <- ctm2 %>% 
  group_by(Site) %>% 
  summarize(mean.CTM = mean(CTM, na.rm = T),
            sd.CTM = sd(CTM, na.rm = T))

ctm.sum

```

New regression with 2019 and 2023 data

```{T  2023}
#call in all CTMs
reg<- read.csv("CTM Regression 2023.csv")

#plot
 ggplot(reg, aes(x = acc.temp, y = mean.ctm, color = loc))+
  geom_smooth(method = "lm", se = F,fullrange=F, size=.5)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = mean.ctm - temp.sd, ymax = mean.ctm + temp.sd), width = 0.2)+
    scale_color_manual("Study Location",
                    breaks = c("Field", "Lab"),
                    labels = c("Field", "Lab"),
                    values = c("dodgerblue", "tomato"))+
  geom_point(x = 17, y = 25.4, size = 2.5, shape=1,color = "tomato")+
  geom_errorbar(aes(x = 17, ymin = 25.4 - 1.7, ymax = 25.4 + 1.7), color = "tomato", width = 0.2)+
  geom_point(x = 16.5, y = 26.5, size = 2.2,shape= 15,color = "dodgerblue")+
  geom_errorbar(aes(x = 16.5, ymin = 26.5 - 0.86, ymax = 26.5 + 0.86), color = "dodgerblue", width = 0.2)+
  scale_y_continuous(breaks = round(seq(min(23), max(30), by = 1.0),1), name="Critical Thermal Maximum (°C)")+
  scale_x_continuous(breaks = round(seq(min(12), max(26), by = 2.0),1), name="Acclimation Temperature (°C)")+
  annotate("text", x = 21.5, y = 24, label = "R^2", parse = T)+
  annotate("text", x = 21.5, y = 23.5, label = " P", parse=T)+
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 12,colour = "black"),
        axis.title.x = element_text(size = 14,colour = "black"),
        axis.title.y = element_text(size = 14,colour = "black"), legend.position=c(0.15,0.85))

 #single regression model by temp and study
modr <- lm(mean.ctm ~ acc.temp*loc, data=reg)
summary(modr)
anova(modr)

#separate regression models
field23 <- reg %>% 
  filter(loc == "Field")

lab23 <- reg %>% 
  filter(loc == "Lab")

field.mod23 <- lm(mean.ctm ~ acc.temp, field23)
anova(field.mod23)
summary(field.mod23)

lab.mod23 <- lm(mean.ctm ~ acc.temp, lab23)
anova(lab.mod23)
summary(lab.mod23)

```